---
title: "Compile data"
author: "Robin Pranter"
date: "2025-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
options(future.globals.maxSize = 120000 * 1024^2) # Change this to fit your 
Sys.setenv(R_MAX_VSIZE = "120Gb")                 # computer

rm(list= ls()); gc()
start_time <- Sys.time()

path <- "~/WorkDirs/PmurNCCPilot_10X_rerun/"

# Load libraries
library(Seurat, lib.loc = "/usr/local/lib/R/site-library")   # Change this to
library(tidyverse, lib.loc = "/usr/local/lib/R/site-library")# mach the location
                                                             # of your libraries
```

#Prep seurat object
Here I will 1. read the dge expression matrices, 2. create one seurat object per 
expression matrix, 3. add meta data, 4. merge all seurat objects into one and 5. 
further add some meta data.


## Read expression data
Here I read in the P30915_109.dgecounts.rds files for each plate. The files 
contain multiple matrices with eg UMI or read counts, exon or intron counts etc.
This makes them large and by by only reading in one of the multiple matrices I 
save a lot of RAM. Be mindful about which matrix to read!

Now I read all inex (intron + exon) mapped UMI-counts.
```{r}
#Read data
raw_counts_P05 <- readRDS(paste0(path, 
              "01_Data/P30915_109/P30915_109.dgecounts.rds"))$umicount$inex$all
raw_counts_P06 <- readRDS(paste0(path, 
              "01_Data/P30915_111/P30915_111.dgecounts.rds"))$umicount$inex$all
```

## Create Seurat objects
```{r}
# Separate gene and ERCC spike-in reads
ercc_counts_P05 <- raw_counts_P05[ grepl("^ERCC", rownames(raw_counts_P05)),]
gene_counts_P05 <- raw_counts_P05[!grepl("^ERCC", rownames(raw_counts_P05)),]

ercc_counts_P06 <- raw_counts_P06[ grepl("^ERCC", rownames(raw_counts_P06)),]
gene_counts_P06 <- raw_counts_P06[!grepl("^ERCC", rownames(raw_counts_P06)),]

#Remove spacious objects
rm(raw_counts_P05,raw_counts_P06); gc()

# Create seurat objects
plate09_SO <- CreateSeuratObject(counts = gene_counts_P05, project = "Plate_09")
plate11_SO <- CreateSeuratObject(counts = gene_counts_P06, project = "Plate_11")

#Remove spacious objects
rm(gene_counts_P05, gene_counts_P06); gc()

# Add a secondary assay with the ERCC reads
plate09_SO[["ERCC"]] = CreateAssayObject(counts = ercc_counts_P05)
plate11_SO[["ERCC"]] = CreateAssayObject(counts = ercc_counts_P06)

rm(ercc_counts_P05, ercc_counts_P06); gc()
```


## Add meta data
### Plate ID
```{r}
# Add the wetlab plate IDs to each seurat object
plates <- c("P30915_109", "P30915_111")

for (i in 1:length(plates)) {
  plate_id <- plates[i]
  #plate_so_name <- paste0("plate", sprintf("%02d", i), "_SO")
  plate_so_name <- c("plate09_SO", "plate11_SO")[i]
  plate_so <- get(plate_so_name)
  plate_so$plateID <- plate_id
  assign(plate_so_name, plate_so)
}
rm(plate_id, plate_so_name, plate_so); gc()
```

### Well ID & exp metadata
The experimental metadata includes embryo ID, phenotype of the parents and 
somite stage.
```{r}
for (i in 1:length(plates)) {
  plate_id <- plates[i]  # Get the plate ID string
  
  # Read the well barcode file
  temp_path <- paste0(path, "01_Data/", plate_id,"/", plate_id, ".well_barcodes.txt")
  bc_well_df <- read.table(temp_path, header = FALSE, skip = 2, 
                           col.names = c("BC", "WellID"))
  
  # Read the experimental metadata file
  temp_path <- paste0(path, "01_Data/", plate_id,"/", plate_id, "_exp_metadata.csv")
  exp_meta_df <- read.table(temp_path, header = TRUE, sep = ",") %>%
    select(c(sampleID, WellID, Pheno, SomSt, FACS_d)) 
    #Some (but not all) metadata files have additional columns with flow cyt 
    #data. This row removes those.
  
  # Merge the well ID df and the experimental metadata df
  merged_df <- merge(bc_well_df, exp_meta_df, by = "WellID")
  
  # Set rownames (needed for AddMetaData()) and remove BC-column
  rownames(merged_df) <- merged_df$BC
  merged_df$BC <- NULL
  
  # Create the Seurat object variable name dynamically
  #plate_so_name <- paste0("plate", sprintf("%02d", i), "_SO")
  plate_so_name <- c("plate09_SO", "plate11_SO")[i] #Simpler since only 2 plates
  
  # Assign metadata to the Seurat object (get() calls an object using a string)
  plate_so <- AddMetaData(get(plate_so_name), metadata = merged_df)
  
  # Assign the temporary seurat objects to permanent ones dynamically
  assign(plate_so_name, plate_so)
}
rm(i, plate_id, temp_path, bc_well_df, exp_meta_df, 
   merged_df, plate_so_name, plate_so); gc()

# The metadata files are written manually following the notes in the lab book. 
# In some simple cases (where all wells got cells from the same embryo) it was written 
#like this:
#
#x <- c()
#for (i in 1:16){
#     for (j in 1:24){
#         x <- append(x, paste0(LETTERS[i], j))
#     }
# }
#
#as.data.frame(list(x,
#                    c(rep("SMA32_C1_E4_2023", 382), "none", "none"), 
#                    c(rep("green", 382), NA, NA), 
#                    c(rep("21ss", 382), NA, NA)), 
#               col.names = c("WellID", "sampleID", "Pheno", "SomSt")) %>% 
#     
#     write.csv(paste0(path, "P309152_107/P309152_107_exp_metadata.csv"), 
#               row.names = FALSE)
#
# FACS date (FACS_d) was later added plate by plate like this:
#read.csv(paste0(path, "P30915_101/P30915_101_exp_metadata.csv")) %>% 
#     mutate(FACS_d = "2024-03-08") %>% 
#     write.csv(paste0(path, "P30915_101/P30915_101_exp_metadata.csv"), 
#               row.names = FALSE)

```

### Read mapping
```{r}
for (i in 1:length(plates)) {
  #plate_name <- paste0("plate", sprintf("%02d", i), "_SO")
  plate_name <- c("plate09_SO", "plate11_SO")[i] #Simpler since only 2 plates
  file_name <- paste0(path, "01_Data/", plates[i], "/", 
                      plates[i], ".readspercell.txt")
  
  # Read and process the reads per cell file
  mapping_perc <- read.table(file_name, header = TRUE) %>%
    group_by(RG) %>%
    summarise(
      total_reads = sum(N),
      mapped_reads = sum(N[type != "Unmapped"], na.rm = TRUE),
      exonic_reads = sum(N[type == "Exon"], na.rm = TRUE),
      intron_reads = sum(N[type == "Intron"], na.rm = TRUE),
      intgen_reads = sum(N[type == "Intergenic"], na.rm = TRUE),
      mapped_perc = (mapped_reads / total_reads) * 100,
      exonic_perc = (exonic_reads / mapped_reads) * 100,
      intron_perc = (intron_reads / mapped_reads) * 100,
      intgen_perc = (intgen_reads / mapped_reads) * 100,
    ) %>%
    as.data.frame()
  # set rownames and remove unwanted columns
  rownames(mapping_perc) <- mapping_perc$RG
  mapping_perc[, c("RG", #"total_reads", 
                   "mapped_reads", "exonic_reads", 
                   "intron_reads", "intgen_reads")] <- NULL 
  
  # Retrieve Seurat object, add metadata, and assign it back
  plate_so <- get(plate_name)
  plate_so <- AddMetaData(plate_so, metadata = mapping_perc)
  assign(plate_name, plate_so)
}
rm(i, plate_name, plate_so, file_name, mapping_perc)
```


## Merege seurat objects
```{r}
all.SO <- merge(plate09_SO, c(plate11_SO), 
                add.cell.ids = c("P30915_109", "P30915_111"))
# Remove known empty wells
all.SO <- subset(all.SO, subset = sampleID != "none")

# Join layers
all.SO[["RNA"]] <- JoinLayers(all.SO[["RNA"]])

#Remove spacious objects
rm(plate09_SO, plate11_SO); gc()

# Inspect the object
all.SO
all.SO@meta.data %>% select(Pheno, SomSt) %>% table()
```

## Further add metadata
### Simplified somite stage
```{r}
# Simlify SomSt
SomSt_Simp <- rep(NA, nrow(all.SO@meta.data))
for (i in 1:nrow(all.SO@meta.data)){
  if (all.SO@meta.data[i, "SomSt"] == "14ss") {
    SomSt_Simp[i] <- "14ss"
  }
  if (all.SO@meta.data[i, "SomSt"] == "15ss"){
    SomSt_Simp[i] <- "14ss"
  }
  if (all.SO@meta.data[i, "SomSt"] == "20ss"){
    SomSt_Simp[i] <- "21ss"
  }
  if (all.SO@meta.data[i, "SomSt"] == "21ss"){
    SomSt_Simp[i] <- "21ss"
  }
  if (all.SO@meta.data[i, "SomSt"] == "28ss"){
    SomSt_Simp[i] <- "28ss"
  }
  if (all.SO@meta.data[i, "SomSt"] == "29ss"){
    SomSt_Simp[i] <- "28ss"
  }
}
all.SO$SomSt_Simp <- SomSt_Simp
rm(SomSt_Simp)

all.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
```

### Calc additional QC meta data
```{r}
## %mitochondrial reads
#Define list of mitochondrial genes
mito_genes <- c("ND1", "CYTB", "ND6", "ND5", "ND4",  #This is all mitochondrial
                "ND4L", "ND3", "COX3", "ATP6",       #genes in mitochondrial 
                "ATP8", "COX2", "COX1", "ND2")       #reference genome 
#NCBI: FJ460597.1
all.SO <- PercentageFeatureSet(all.SO,
                               features = mito_genes,
                               col.name = "percent.mt")

## %ribosomal reads
#Define list of ribosomal genes
ribo_genes <- rownames(all.SO)[grep("^RP[SL]", rownames(all.SO))]
all.SO <- PercentageFeatureSet(all.SO,
                               features = ribo_genes,
                               col.name = "percent.rb")

## Look at the SO before save
names(all.SO@meta.data)
all.SO
```

# Save data
Save the compiled data and meta data in an RDS file to be read by the next script in
the pipeline (02_FilterSmrtS.Rmd)
```{r}
saveRDS(all.SO, paste0(path, 
                       "02_Analysis/01_CompileSmrtS_output_", 
                       Sys.Date(),
                       ".Rds"))
```

Calculated runtime:
```{r}
end_time <- Sys.time()
paste0("Elapsed time: ", end_time - start_time)
sessionInfo()
```