---
title: "filtering"
author: "Robin Pranter"
date: "2025-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
options(future.globals.maxSize = 120000 * 1024^2) # Change this to fit your 
Sys.setenv(R_MAX_VSIZE = "120Gb")                 # computer

rm(list= ls()); gc()
start_time <- Sys.time()

path <- "~/WorkDirs/PmurNCCPilot_10X_rerun/"

# Load libraries
library(Seurat, lib.loc = "/usr/local/lib/R/site-library")   # Change this to
library(tidyverse, lib.loc = "/usr/local/lib/R/site-library")# mach the location
library(patchwork, lib.loc = "/usr/local/lib/R/site-library")# of your libraries
```

# Read data
Read in the RDS-object that has all the data and metadata calculated and 
compiled by 01_CompileSmrtS.Rmd. 
```{r}
filename <- list.files(paste0(path, "02_Analysis/"), 
           pattern = "01_CompileSmrtS_output_")
if (length(filename) == 1) {
  all.SO <- readRDS(file = paste0(path, "02_Analysis/", filename))
  } else {
    print(paste0("You have two files matching the pattern. ", 
                 "Choose one to read and remove the other from the directory."))
    print(filename)
    }
rm(filename)
```

## Basic stats
```{r}
all.SO@meta.data %>% 
  subset(SomSt == "28ss") %>%
  summarize(n = n(),
            av_count = mean(nCount_RNA) %>% round(),
            sd_count = sd(nCount_RNA) %>% round(),
            av_genes = mean(nFeature_RNA) %>% round(),
            sd_genes = sd(nFeature_RNA) %>% round()) %>%
  t()
```

# Filter
```{r}
# Define a list with the cell filter cutoff values
c_filters <- list(nCount_RNA    = 1500,
                  nFeature_RNA  = 1500,
                  nCount_ERCC   = 300,
                  nFeature_ERCC = NA,
                  mapped_perc   = 40,
                  exonic_perc   = 73,
                  intron_perc   = NA,
                  intgen_perc   = NA,
                  percent.mt    = 16,
                  percent.rb    = 7)
# Feture filter
f_filter <- 5
```

## Vissualize filters
### Feature filter
```{r}
P04 <- list(
  Matrix::rowSums(all.SO[["RNA"]]$counts > 0) %>% 
  as.data.frame %>% 
  dplyr::rename(num_cells = ".") %>%
  ggplot(aes(x = num_cells)) +
  #geom_density(fill = "blue", alpha = 0.5) +
  geom_histogram(fill = "blue", alpha = 0.5, bins = 120) +
  labs(title = "Density of detected genes per cell",
       x = "Number of cells a gene is detected in",
       y = "Density") +
  geom_vline(xintercept = f_filter) +
  theme_minimal() + 
  theme(text = element_text(size = 6),
        axis.title.x = element_text(size = 6),
        axis.title.y = element_text(size = 6),
        axis.text = element_text(size = 6, angle = 45),
        legend.position = "none"),
  
  Matrix::rowSums(all.SO[["RNA"]]$counts > 0) %>%
  as.data.frame %>% 
  dplyr::rename(num_cells = ".") %>%
  ggplot(aes(x = num_cells)) +
  #geom_density(fill = "purple", alpha = 0.5) +
  geom_histogram(fill = "purple", alpha = 0.5, bins = 120) +
  scale_x_continuous(transform = "log1p") +
  labs(title = "Density of detected genes per cell",
       x = "Number of cells a gene is detected in (ln(x+1))",
       y = "Density") +
  geom_vline(xintercept = f_filter) +
  theme_minimal() +
  theme(text = element_text(size = 6),
        axis.title.x = element_text(size = 6),
        axis.title.y = element_text(size = 6),
        axis.text = element_text(size = 6, angle = 45),
        legend.position = "none")
)
P04 <- wrap_plots(P04, ncol = 1)
P04
```

```{r}
P05 <- cbind(Matrix::rowSums(all.SO[["RNA"]]$counts > 0),
             Matrix::rowSums(all.SO[["RNA"]]$counts)) %>%
  as.data.frame() %>%
  dplyr::rename(num_cells = "V1", tot_exp = "V2") %>%
  ggplot(aes(x = tot_exp, y = num_cells)) +
  geom_point(alpha = .1, col = "blue") + 
  scale_x_continuous(transform = "log1p") +
  scale_y_continuous(transform = "log1p") +
  geom_hline(yintercept = f_filter) +
  labs(y = "Number of cells a gene is detected in (ln(x+1))",
       x = "total expression (ln(x+1))") +
  theme(text = element_text(size = 6),
        axis.title.x = element_text(size = 6),
        axis.title.y = element_text(size = 6),
        axis.text = element_text(size = 6, angle = 45, hjust = 1))
P05
```

### Cell filters
```{r}
# Create a VlnPlot for each feature
P06 <- lapply(seq_along(c_filters), function(i) {
  p <- VlnPlot(all.SO, features = names(c_filters)[i], 
               pt.size = .1, assay = "RNA", layer = "counts")
  
  # Add geom_hline only if the value is not NA
  if (!is.na(c_filters[[i]])) {
    p <- p + geom_hline(yintercept = c_filters[[i]]) +
      geom_text(aes(x = .2, y = c_filters[[i]], label = c_filters[[i]]), 
                col = "black", hjust = 0, vjust = -0.5, 
                size = 8*.35) # *.35 translates pt to mm
  }
  
  #Make it pretty with themes
  p + theme(text = element_text(size = 6),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 6, angle = 45),
            axis.text.x = element_blank(),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 6),
            legend.key.size = unit(3, "mm"))
  })

# Combine all plots into a single plot
P06 <- wrap_plots(P06, ncol = 5) + 
  plot_layout(guides = "collect") +
  plot_annotation("All data")
# Display the combined plot
P06
```

## Detection based filter
By "detection based filtering" I mean removal of cells with low numbers of 
detected genes, and removal of genes that are detected in few cells.
```{r}
# Filter cells
selected_c <- WhichCells(all.SO, 
                         expression = nFeature_RNA > c_filters$nFeature_RNA)
# Filter features (=genes)
selected_f <- rownames(all.SO)[
  Matrix::rowSums(all.SO[["RNA"]]$counts > 0) >= f_filter]
# This removes all genes that are detected in 4 or fewer cells.

#The actual filtering happens here
det.SO <- subset(all.SO, cells = selected_c, features = selected_f)
rm(selected_c, selected_f)

#Compare the unfiltered and filtered data sets
dim(all.SO)
dim(det.SO)
print(paste0("Genes removed: ", dim(all.SO)[1] - dim(det.SO)[1]))
print(paste0("Cells removed: ", dim(all.SO)[2] - dim(det.SO)[2]))
all.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
det.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
```

## Targeted genes removal 
```{r}
#Remove mitochondrial genes
#Define list of mitochondrial genes
mito_genes <- c("ND1", "CYTB", "ND6", "ND5", "ND4",  #This is all mitochondrial
                "ND4L", "ND3", "COX3", "ATP6",       #genes in mitochondrial 
                "ATP8", "COX2", "COX1", "ND2")       #reference genome 
selected_f <- rownames(det.SO)[!(rownames(det.SO) %in% mito_genes)]

#Remove ribosomal genes
selected_f <- selected_f[!grepl("^RP[SL]", selected_f)] # This removes rRNA

# Remove all genes named UnnamedGeneX
selected_f <- selected_f[!grepl("UnnamedGene", selected_f)]
#By running 
#"grep 'UnnamedGene' GCF_004329235.1_PodMur_1.0_genomic_mod.gtf | less -S"
#I could see that these were all tRNA or rRNA

# Filter
det.SO <- subset(det.SO, features = selected_f)
rm(selected_f, mito_genes); gc()

#Compare the unfiltered and filtered data sets
dim(all.SO)
dim(det.SO)
print(paste0("Genes removed: ", dim(all.SO)[1] - dim(det.SO)[1]))
print(paste0("Cells removed: ", dim(all.SO)[2] - dim(det.SO)[2]))
all.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
det.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
```

### Plot QCs after detection based filt
```{r}
# Create a VlnPlot for each feature
P07 <- lapply(seq_along(c_filters), function(i) {
  p <- VlnPlot(det.SO, features = names(c_filters)[i], 
               pt.size = .1, assay = "RNA", layer = "counts")
  
  # Add geom_hline only if the value is not NA
  if (!is.na(c_filters[[i]]) & names(c_filters)[i] != "nFeature_RNA") {
    p <- p + geom_hline(yintercept = c_filters[[i]]) +
      geom_text(aes(x = .2, y = c_filters[[i]], label = c_filters[[i]]), 
                col = "black", hjust = 0, vjust = -0.5, 
                size = 8*.35) # *.35 translates pt to mm
  }
  
  #Make it pretty with themes
  p + theme(text = element_text(size = 6),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 6, angle = 45),
            axis.text.x = element_blank(),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 6),
            legend.key.size = unit(3, "mm"))
  })

# Combine all plots into a single plot
P07 <- wrap_plots(P07, ncol = 5) + 
  plot_layout(guides = "collect") +
  plot_annotation("Detection filtered data")
# Display the combined plot
P07
```

By only filtering based on nFeature_RNA most of the potential problems in the 
data is already removed. Here I mainly mean the hump of cells with low values in nCount_RNA, mapped_perc and exonic_pers. This is good, it means that the QCs 
tell a coherent story. I will do some additional filtering of the other QCs.

## QC-metric filtering
```{r}
# Select cells
selected_nCou <- WhichCells(det.SO,expression=nCount_RNA >c_filters$nCount_RNA)
selected_mito <- WhichCells(det.SO,expression=percent.mt <c_filters$percent.mt)
selected_mapp <- WhichCells(det.SO,expression=mapped_perc>c_filters$mapped_perc)
selected_exon <- WhichCells(det.SO,expression=exonic_perc>c_filters$exonic_perc)
selected_ERCC <- WhichCells(det.SO,expression=nCount_ERCC<c_filters$nCount_ERCC)
selected_ribo <- WhichCells(det.SO,expression=percent.rb >c_filters$percent.rb)

#and subset to only keep those cells
filt.SO <- subset(det.SO, cells = Reduce(intersect, list(selected_nCou,
                                                         selected_mito,
                                                         selected_mapp,
                                                         selected_exon,
                                                         selected_ERCC,
                                                         selected_ribo)))
#Compare the unfiltered and filtered data sets
dim(all.SO)
dim(det.SO)
dim(filt.SO)
print(paste0("Genes removed (det based filt): ",dim(all.SO)[1]-dim(det.SO)[1]))
print(paste0("Cells removed (det based filt): ",dim(all.SO)[2]-dim(det.SO)[2]))
print(paste0("Add. cells removed (full filt): ",dim(det.SO)[2]-dim(filt.SO)[2]))
all.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
det.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
filt.SO@meta.data %>% select(Pheno, SomSt_Simp) %>% table()
```

### Plot QCs after QC based filt
```{r}
# Create a VlnPlot for each feature
P08 <- lapply(seq_along(c_filters), function(i) {
  p <- VlnPlot(filt.SO, features = names(c_filters)[i], 
               pt.size = .1, assay = "RNA", layer = "counts")
  
  # Add geom_hline only if the value is not NA
  if (!is.na(c_filters[[i]]) & names(c_filters)[i] != "nFeature_RNA") {
    p <- p + geom_hline(yintercept = c_filters[[i]]) +
      geom_text(aes(x = .2, y = c_filters[[i]], label = c_filters[[i]]), 
                col = "black", hjust = 0, vjust = -0.5, 
                size = 8*.35) # *.35 translates pt to mm
  }
  
  #Make it pretty with themes
  p + theme(text = element_text(size = 6),
            axis.title.x = element_blank(),
            axis.text.y = element_text(size = 6, angle = 45),
            axis.text.x = element_blank(),
            legend.text = element_text(size = 6),
            legend.title = element_text(size = 6),
            legend.key.size = unit(3, "mm"))
  })

# Combine all plots into a single plot
P08 <- wrap_plots(P08, ncol = 5) + 
  plot_layout(guides = "collect") +
  plot_annotation("QC filtered data")
# Display the combined plot
P08
```

# Save data
Save the filtered data and meta data in an RDS file to be read by the next script in the pipeline (04_IntAnalysis10XAndSmrtS.Rmd)
```{r}
filt.SO
saveRDS(filt.SO, paste0(path, 
                       "02_Analysis/02_FilterSmrtS_output_", 
                       Sys.Date(),
                       ".Rds"))
```

Calculated runtime:
```{r}
end_time <- Sys.time()
paste0("Elapsed time: ", end_time - start_time)
sessionInfo()
```
