---
title: "Untitled"
output: html_document
date: "2026-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up 

Set up workspace, read libraries and source functions. 
```{r, message=FALSE, echo=FALSE}
# Increase maximum memory available to R
options(future.globals.maxSize = 120000 * 1024^2) # Change this to fit your 
Sys.setenv(R_MAX_VSIZE = "120Gb")                 # computer

rm(list = ls()); invisible(gc())
start_time <- Sys.time()

# Load libraries
library(Seurat, lib.loc = "/usr/local/lib/R/site-library")   # Change this to
library(tidyverse, lib.loc = "/usr/local/lib/R/site-library")# mach the location
library(DoubletFinder)                                       # of your libraries

path <- "~/WorkDirs/PmurNCCPilot_10X_rerun/"
```

# Read data
```{r}
#Load data
Pmur_82.SO <- Seurat::Read10X_h5(filename = paste0(path,
                   "01_Data/10X_21_082_P22202_8007_filtered_feature_bc_matrix.h5"),
                    use.names = T) %>%
  CreateSeuratObject(project = "Pmur_82")

Pmur_84.SO <- Seurat::Read10X_h5(filename = paste0(path, 
                   "01_Data/10X_21_084_P22202_8008_filtered_feature_bc_matrix.h5"),
                    use.names = T) %>%
  CreateSeuratObject(project = "Pmur_84")

Pmur_85.SO <- Seurat::Read10X_h5(filename = paste0(path, 
                   "01_Data/10X_21_085_P22202_8009_filtered_feature_bc_matrix.h5"),
                    use.names = T) %>%
  CreateSeuratObject(project = "Pmur_85")

#Add Sorted/unsorted in metadata
Pmur_82.SO$type <- "Sorted"
Pmur_84.SO$type <- "Unsort"
Pmur_85.SO$type <- "Unsort"

# Merge objects
all.SO <- merge(Pmur_82.SO, c(Pmur_84.SO, Pmur_85.SO), 
                add.cell.ids = c("Pmur_82.SO", "Pmur_84.SO", "Pmur_85.SO"))
rm(Pmur_82.SO, Pmur_84.SO, Pmur_85.SO); invisible(gc())

# Join layers
all.SO[["RNA"]] <- JoinLayers(all.SO[["RNA"]])
# (the counts layers in the RNA assays gets added as separate layers for each
# sample in the merged object. Here I join them together in one layer)

all.SO
```

## Basic stats
```{r}
all.SO@meta.data %>%
  group_by(type) %>% 
  summarize(n = n(),
            av_count = mean(nCount_RNA) %>% round(),
            sd_count = sd(nCount_RNA) %>% round(),
            av_genes = mean(nFeature_RNA) %>% round(),
            sd_genes = sd(nFeature_RNA) %>% round()) %>%
  t()
```


# Calc metadata
```{r}
## %mitochondrial reads
#Define list of mitochondrial genes
mito_genes <- c("ND1", "CYTB", "ND6", "ND5", "ND4",  #This is all mitochondrial
                "ND4L", "ND3", "COX3", "ATP6",       #genes in mitochondrial 
                "ATP8", "COX2", "COX1", "ND2")       #reference genome 
#NCBI: FJ460597.1
all.SO <- PercentageFeatureSet(all.SO,
                               features = mito_genes,
                               col.name = "percent.mt")

## %ribosomal reads
#Define list of ribosomal genes
ribo_genes <- rownames(all.SO)[grep("^RP[SL]", rownames(all.SO))]
all.SO <- PercentageFeatureSet(all.SO,
                               features = ribo_genes,
                               col.name = "percent.rb")

## list meta data variables
names(all.SO@meta.data)
```

# Filter
I do three types of filtering: detection based filtering, filtering based on 
QC-metrics and removal of doublets.

## Detection based filtering
By "detection based filtering" I mean removal of cells with low numbers of 
detected genes and removal of genes that are detected in few cells.
```{r}
#Select cells
selected_c <- WhichCells(all.SO, expression = nFeature_RNA > 700)

#Select genes
#selected_f <- rownames(all.SO)[
#  Matrix::rowSums(all.SO@assays$RNA@layers$counts) > 3]
selected_f <- rownames(all.SO)[
  Matrix::rowSums(all.SO[["RNA"]]$counts > 0) >= 20]

#Filter
filt.SO <- subset(all.SO, cells = selected_c, features = selected_f)
rm(selected_c, selected_f)

#Also remove the extreme outlyer in nCount_RNA.
selected_c <- WhichCells(all.SO, expression = nCount_RNA < 90000)
filt.SO <- subset(filt.SO, cells = selected_c)
rm(selected_c)

#also remove ribo and mito genes
filt.SO <- filt.SO[ ! rownames(filt.SO) %in% mito_genes, ]
filt.SO <- filt.SO[ ! rownames(filt.SO) %in% ribo_genes, ]

dim(all.SO)
dim(filt.SO)
table(all.SO$orig.ident)
table(filt.SO$orig.ident)
#rm(all.SO);invisible(gc())
```

## QC-metric filtering
```{r}
# Select cells
selected_mito <- WhichCells(filt.SO, expression = percent.mt < 15)
selected_ribo <- WhichCells(filt.SO, expression = percent.rb > 15)

#and subset to only keep those cells
filt.SO <- subset(filt.SO, cells = selected_mito)
filt.SO <- subset(filt.SO, cells = selected_ribo)
rm(selected_mito, selected_ribo)

dim(filt.SO)
```

## Cell cykle state
Since the sorted sample is sorted based on DAPI while the unsorted samples are
unsorted it is important to check that the samples don't differ in cell cycle.

### Prep SO for CC-scoring
```{r}
# Normalize and scale data and find variable features (all in SCTransform())
filt.SO <- SCTransform(filt.SO,
                       assay = 'RNA',
                       new.assay.name = 'SCT',
                       vars.to.regress = c('percent.mt'),
                       method = "glmGamPoi",
                       variable.features.n = 3000) #3000 is default
```

### Find CC-markers
#### Read Lizard Human Ortho-table
```{r}
Ortho.df <- read.csv(file=paste0(path, "01_Data/",
                          "Pmur1.0Human_OrthoTable_BioMartManDL20241206.txt"))
```

#### Find Lizard S and G2M orthos
```{r}
# Save S and G2M markers in Pmur
G2M.genes <- Ortho.df[Ortho.df$Human.gene.name %in% 
                        cc.genes.updated.2019$g2m.genes, ]$Gene.name
S.genes <- Ortho.df[Ortho.df$Human.gene.name %in% 
                      cc.genes.updated.2019$s.genes, ]$Gene.name

#Check that the markers are present in the data
print(paste0(filt.SO %>% Features() %in% G2M.genes %>% sum(), "/",
             G2M.genes %>% length(), 
             " G2M-markers are present in the data"))
print(paste0(filt.SO %>% Features() %in% S.genes %>% sum(), "/",
             S.genes %>% length(), 
             " S-markers are present in the data"))

#Subset only the genes that are present
G2M.genes <- G2M.genes[G2M.genes %in% Features(filt.SO)]
S.genes <- S.genes[S.genes %in% Features(filt.SO)]
```

### Calc score
```{r}
# Predict cell cycle state (make sure to specify the "assay" parameter)
filt.SO <- CellCycleScoring(filt.SO,
                            s.features = S.genes,
                            g2m.features = G2M.genes,
                            assay = 'SCT',
                            set.ident = TRUE)
```

## Doublet filter
To remove doublet cells I will use the package doublet finder (DF). First I need
a UMAP and cluster classifications. Then I can estimate some parameters for the
doublet prediction. Last I can run the prediction and remove the predicted
doublets. One of the parameters (the estimated doublet percentage, here names
EstDoubPerc) can not be estimated from the data but must be set by myself 
informed by the literature. 

### Prep SO for DF
Doublet finder needs clusters to calc doublets. Here I run the standard Seurat 
pipeline for clustering. I will redo this later after all filtering is finished. 
```{r}
#Reduce dimensions using PCA and UMAP
filt.SO <- RunPCA(filt.SO, verbose = F, npcs = 20, assay = "SCT") 
#Cluster
filt.SO <- FindNeighbors(filt.SO, dims = 1:10)
filt.SO <- FindClusters(filt.SO)

#DF should be run on each sample separately. 
#Therefor I split the SO in to six SOs and store them in a list
HVGs <- VariableFeatures(filt.SO)
split.list <- SplitObject(filt.SO, split.by = "orig.ident")
```

### Predict doublets
#### Est pK & set exp % doublets
```{r}
#pK Identification (no ground-truth)
#sweep.list <- paramSweep(split.list$Pmur_85, PCs = 1:10, sct = TRUE) #choose sample
#sweep.stat <- summarizeSweep(sweep.list, GT = FALSE)
#bcmvn_Pmur <- find.pK(sweep.stat)

#Plot pK values against the mean-variance normalized bimodality coefficient
#plot(as.numeric(as.character(bcmvn_Pmur$pK)), bcmvn_Pmur$BCmetric, type = "b")
#subset(bcmvn_Pmur, subset = BCmetric == max(bcmvn_Pmur$BCmetric))

#Set estimated pK based on plot from find.pK() above 
est_pK_82 <- .07
est_pK_84 <- .05
est_pK_85 <- .09

#Set exp perc doublets based on table under point 8 here: 
#https://nbisweden.github.io/workshop-scRNAseq/labs/seurat/seurat_01_qc.html
ExpDubPerc_82 <- .03
ExpDubPerc_84 <- .065
ExpDubPerc_85 <- .05

#Based on expected doublet percentage, Homotypic Doublet Proportion Estimate
homotypic.prop_82 <-modelHomotypic(split.list$Pmur_82@meta.data$seurat_clusters)
nExp_poi_82       <-round(ExpDubPerc_82*nrow(split.list$Pmur_82@meta.data))

homotypic.prop_84 <-modelHomotypic(split.list$Pmur_84@meta.data$seurat_clusters)
nExp_poi_84       <-round(ExpDubPerc_84*nrow(split.list$Pmur_84@meta.data))

homotypic.prop_85 <-modelHomotypic(split.list$Pmur_85@meta.data$seurat_clusters)
nExp_poi_85       <-round(ExpDubPerc_85*nrow(split.list$Pmur_85@meta.data))
```

#### Run DoubletFinder
```{r, message = FALSE}
#Using the unadjusted poison expected doublet number -nExp_poi_Px-
split.list$Pmur_82 <- doubletFinder(split.list$Pmur_82, 
                                    PCs = 1:10,
                                    pN = 0.25, 
                                    pK = est_pK_82, 
                                    nExp = nExp_poi_82, 
                                    reuse.pANN = NULL, 
                                    sct = TRUE)
names(split.list$Pmur_82@meta.data)[grepl("DF.classification", 
                              names(split.list$Pmur_82@meta.data))] <- "DF.pred"
sum(split.list$Pmur_82@meta.data$DF.pred == "Singlet")
nrow(split.list$Pmur_82@meta.data)

split.list$Pmur_84 <- doubletFinder(split.list$Pmur_84, 
                                    PCs = 1:10,
                                    pN = 0.25, 
                                    pK = est_pK_84, 
                                    nExp = nExp_poi_84, 
                                    reuse.pANN = NULL, 
                                    sct = TRUE)
names(split.list$Pmur_84@meta.data)[grepl("DF.classification", 
                              names(split.list$Pmur_84@meta.data))] <- "DF.pred"
sum(split.list$Pmur_84@meta.data$DF.pred == "Singlet")
nrow(split.list$Pmur_84@meta.data)

split.list$Pmur_85 <- doubletFinder(split.list$Pmur_85, 
                                    PCs = 1:10,
                                    pN = 0.25, 
                                    pK = est_pK_85, 
                                    nExp = nExp_poi_85, 
                                    reuse.pANN = NULL, 
                                    sct = TRUE)
names(split.list$Pmur_85@meta.data)[grepl("DF.classification", 
                              names(split.list$Pmur_85@meta.data))] <- "DF.pred"
sum(split.list$Pmur_85@meta.data$DF.pred == "Singlet")
nrow(split.list$Pmur_85@meta.data)
gc()
```

### Merge data
```{r}
#Merge
filt.SO <- merge(split.list$Pmur_82, c(split.list$Pmur_84, split.list$Pmur_85))
filt.SO[["RNA"]] <- JoinLayers(filt.SO[["RNA"]])
VariableFeatures(filt.SO) <- HVGs; rm(HVGs)
```

### Remove doublets
Judgeing by the nCount_RNA, the predicted doublets might not even be doublets 
(they don't have higher than expected gene expression). However, I will remove 
them anyway to be conservative.

```{r}
#Filter
dim(filt.SO)
filt.SO <- subset(filt.SO, DF.pred == "Singlet")
dim(filt.SO)

#Remove spacious objects
rm(split.list,
   est_pK_82, est_pK_84, est_pK_85,
   
   homotypic.prop_82, homotypic.prop_84, homotypic.prop_85, 
   
   nExp_poi.adj_82, nExp_poi.adj_84, nExp_poi.adj_85,
   
   nExp_poi_82, nExp_poi_84, nExp_poi_85,
   
   ExpDubPerc_82, ExpDubPerc_84, ExpDubPerc_85); invisible(gc())
```

## Filter cells with high exp
```{r}
#select cells to keep
selected_Count <- WhichCells(filt.SO, expression = nCount_RNA < 25000)
#and subset to only keep those cells
filt.SO <- subset(filt.SO, cells = selected_Count)
rm(selected_Count)
dim(filt.SO)
```


```{r}
VlnPlot(filt.SO, c("nCount_RNA", "nFeature_RNA", "percent.mt", 
                   "percent.rb", "S.Score", "G2M.Score"),
        group.by = "orig.ident", pt.size = 0)
```


# Check QCs in reduced dims 
Check that QCs don't distribute unevenly in reduced expression spaces.

## Dimensionality reduction
```{r}
#Reset default assay to the raw RNA counts
DefaultAssay(filt.SO) <- "RNA"

# Normalize, transform and select variable features
filt.SO <- SCTransform(filt.SO,                 
                       new.assay.name = "filt_SCT_Reg",       
                       vars.to.regress =c("percent.mt", "G2M.Score", "S.Score"),
                       method = "glmGamPoi",            
                       variable.features.n = 3000) #3000 is default

#Reduce dimensions
filt.SO <- RunPCA(filt.SO) %>%
  RunUMAP(reduction = "pca", dims = 1:10)
```

```{r}
FeaturePlot(filt.SO, c("nCount_RNA", "nFeature_RNA", "percent.mt",
                       "percent.rb", "S.Score", "G2M.Score"))
```

# Save data
Save RDS to continue with the second half of the analysis in a separate script
```{r}
saveRDS(filt.SO, paste0(path, "02_Data/",
                        "03_CompileAndFilter10X_out_", 
                        Sys.Date(),
                        ".Rds"))
filt.SO
```

Calculated runtime:
```{r}
end_time <- Sys.time()
paste0("Elapsed time: ", end_time - start_time)
sessionInfo()
```
