---
title: "Untitled"
author: "Robin Pranter"
date: "2025-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up
```{r}
options(future.globals.maxSize = 120000 * 1024^2) # Change this to fit your 
Sys.setenv(R_MAX_VSIZE = "120Gb")                 # computer

rm(list=ls());gc()
start_time <- Sys.time()

library(dplyr, lib.loc = "/usr/local/lib/R/site-library")   # Change this to
library(Seurat, lib.loc = "/usr/local/lib/R/site-library")  # mach the location
library(ggplot2, lib.loc = "/usr/local/lib/R/site-library") # of your libraries
library(reshape2, lib.loc = "/usr/local/lib/R/site-library")
library(ggridges, lib.loc = "/usr/local/lib/R/site-library")
library(ggrastr, lib.loc = "/usr/local/lib/R/site-library")

path <- "~/WorkDirs/PmurNCCPilot_10X_rerun/"
```

# Read Soldatov data
```{r}
#Read count tables
Sol_1 <- read.table(paste0(path, "01_Data/SoldatovData/",
                           "E10.5_post-otic_Sox10_counts.txt"), sep = " ")
Sol_1$gene <- rownames(Sol_1)
Sol_2 <- read.table(paste0(path, "01_Data/SoldatovData/",
                           "E10.5_tail_Wnt1_counts.txt"), sep = " ")
Sol_2$gene <- rownames(Sol_2)
Sol_3 <- read.table(paste0(path, "01_Data/SoldatovData/",
                           "E10.5_trunk_Wnt1_counts.txt"), sep = " ")
Sol_3$gene <- rownames(Sol_3)
Sol_4 <- read.table(paste0(path, "01_Data/SoldatovData/", 
                           "E8.5_whole_embryo_Wnt1_counts.txt"), sep = " ")
Sol_4$gene <- rownames(Sol_4)
Sol_5 <- read.table(paste0(path, "01_Data/SoldatovData/",
                           "E9.5_anterior_Sox10_counts.txt"), sep = " ")
Sol_5$gene <- rownames(Sol_5)
Sol_6 <- read.table(paste0(path, "01_Data/SoldatovData/",
                           "E9.5_posterior_Sox10_counts.txt"), sep = " ")
Sol_6$gene <- rownames(Sol_6)
Sol_7 <- read.table(paste0(path, "01_Data/SoldatovData/",
                           "E9.5_trunk_Wnt1_counts.txt"), sep = " ")
Sol_7$gene <- rownames(Sol_7)

#Merge tables into one
Sol <- merge(Sol_1, Sol_2, by = "gene", all.x = TRUE, all.y = TRUE) %>%
  merge(Sol_3, by = "gene", all.x = TRUE, all.y = TRUE) %>%
  merge(Sol_4, by = "gene", all.x = TRUE, all.y = TRUE) %>%
  merge(Sol_5, by = "gene", all.x = TRUE, all.y = TRUE) %>%
  merge(Sol_6, by = "gene", all.x = TRUE, all.y = TRUE) %>%
  merge(Sol_7, by = "gene", all.x = TRUE, all.y = TRUE)
#Remove old tables
rm(Sol_1, Sol_2, Sol_3, Sol_4, Sol_5, Sol_6, Sol_7); gc()

#Make nice
Sol[,-1] <- lapply(Sol[,-1], as.numeric)
Sol <- Sol[order(colnames(Sol))]

#Use gene names for rownames and remove gene column
rownames(Sol) <- Sol$gene
Sol <- dplyr::select(Sol, -c("gene"))
gc()
```

```{r}
# Extract the row and convert to long format
rel_exp.df <- data.frame(Sox10_PM = as.numeric(Sol["Sox10", ]/
                                                    (colSums(Sol, na.rm = TRUE)/
                                                       1e+06)),
                         Sox9_PM = as.numeric(Sol["Sox9", ]/
                                                    (colSums(Sol, na.rm = TRUE)/
                                                       1e+06)),
                         Snai1_PM = as.numeric(Sol["Snai1", ]/
                                                    (colSums(Sol, na.rm = TRUE)/
                                                       1e+06)),
                         Snai2_PM = as.numeric(Sol["Snai2", ]/
                                                    (colSums(Sol, na.rm = TRUE)/
                                                       1e+06)),
                         Tfap2a_PM = as.numeric(Sol["Tfap2a", ]/
                                                   (colSums(Sol, na.rm = TRUE)/
                                                       1e+06)),
                            Dataset = "Soldatov",
                            Sorted = "Sox10")
rel_exp.df[is.na(rel_exp.df)] <- 0 # Some genes are not present in the count 
#rm(Sol); gc()                     # table, introducing NAs. These are converted
                                   # to 0s here
```

# Read Cao data
```{r}
#Read data
rel_exp.df <- readRDS(paste0(path, 
                      "01_Data/CaoData/gene_count_cleaned.RDS")) %>%
  CreateSeuratObject(project = "Cao2019") %>%
  FetchData(vars = c("ENSMUSG00000033006.9",  #Sox10
                     "ENSMUSG00000000567.5",  #Sox9
                     "ENSMUSG00000042821.7",  #Snai1
                     "ENSMUSG00000022676.6",  #Snai2
                     "ENSMUSG00000021359.15", #Tfap2a
                     "nCount_RNA"), layer = "counts") %>%
  mutate(Sox10_PM  = ENSMUSG00000033006.9  / (nCount_RNA / 1e+06),
         Sox9_PM   = ENSMUSG00000000567.5  / (nCount_RNA / 1e+06),
         Snai1_PM  = ENSMUSG00000042821.7  / (nCount_RNA / 1e+06),
         Snai2_PM  = ENSMUSG00000022676.6  / (nCount_RNA / 1e+06),
         Tfap2a_PM = ENSMUSG00000021359.15 / (nCount_RNA / 1e+06),
         Dataset = "Cao",
         Sorted = "Whole embryo") %>%
  select(c("Sox10_PM", "Sox9_PM", "Snai1_PM", "Snai2_PM", 
           "Tfap2a_PM", "Dataset", "Sorted")) %>%
  rbind(rel_exp.df)
gc()
```

# Read 10X data
```{r}
filename10X <- list.files(paste0(path, "02_Analysis/"), 
           pattern = "03_CompileAndFilter10X_out_")
if (length(filename10X) > 1) {
  print(paste0("You have two files matching the pattern.", 
               "Choose one to read and remove the other from the directory."))
  print(filename10X)
}
```

```{r}
rel_exp.df <- readRDS(paste0(path, "02_Analysis/", filename10X)) %>%
  FetchData(vars = c("SOX10",
                     "SOX9",
                     "SNAI1", 
                     "SNAI2",
                     "TFAP2A",
                     "nCount_RNA", 
                     "orig.ident"), layer = "counts") %>%
  subset(orig.ident == "Pmur_82") %>%
  mutate(Sox10_PM = SOX10 / (nCount_RNA / 1e+06),
         Sox9_PM  = SOX9 / (nCount_RNA / 1e+06),
         Snai1_PM = SNAI1 / (nCount_RNA / 1e+06),
         Snai2_PM = SNAI2 / (nCount_RNA / 1e+06),
         Tfap2a_PM = TFAP2A / (nCount_RNA / 1e+06),
         Dataset = "Leniently sorted",
         Sorted = "HNK-1 top 15%") %>%
  select(c("Sox10_PM", "Sox9_PM", "Snai1_PM", "Snai2_PM", 
           "Tfap2a_PM", "Dataset", "Sorted")) %>%
  rbind(rel_exp.df)

rel_exp.df <- readRDS(paste0(path, "02_Analysis/", filename10X)) %>%
  FetchData(vars = c("SOX10",
                     "SOX9",
                     "SNAI1", 
                     "SNAI2",
                     "TFAP2A",
                     "nCount_RNA", 
                     "orig.ident"), layer = "counts") %>%
  subset(orig.ident == "Pmur_84") %>% 
  mutate(Sox10_PM = SOX10 / (nCount_RNA / 1e+06),
         Sox9_PM  = SOX9 / (nCount_RNA / 1e+06),
         Snai1_PM = SNAI1 / (nCount_RNA / 1e+06),
         Snai2_PM = SNAI2 / (nCount_RNA / 1e+06),
         Tfap2a_PM = TFAP2A / (nCount_RNA / 1e+06),
         Dataset = "Unsorted",
         Sorted = "Whole embryo") %>%
  select(c("Sox10_PM", "Sox9_PM", "Snai1_PM", "Snai2_PM", 
           "Tfap2a_PM", "Dataset", "Sorted")) %>%
  rbind(rel_exp.df)

rel_exp.df <- readRDS(paste0(path, "02_Analysis/", filename10X)) %>%
  FetchData(vars = c("SOX10",
                     "SOX9",
                     "SNAI1",
                     "SNAI2",
                     "TFAP2A",
                     "nCount_RNA", 
                     "orig.ident"), layer = "counts") %>%
  subset(orig.ident == "Pmur_85") %>% 
  mutate(Sox10_PM = SOX10 / (nCount_RNA / 1e+06),
         Sox9_PM  = SOX9 / (nCount_RNA / 1e+06),
         Snai1_PM = SNAI1 / (nCount_RNA / 1e+06),
         Snai2_PM = SNAI2 / (nCount_RNA / 1e+06),
         Tfap2a_PM = TFAP2A / (nCount_RNA / 1e+06),
         Dataset = "Unsorted",
         Sorted = "Whole embryo") %>%
  select(c("Sox10_PM", "Sox9_PM", "Snai1_PM", "Snai2_PM", 
           "Tfap2a_PM", "Dataset", "Sorted")) %>%
  rbind(rel_exp.df)
rm(filename10X); gc()
```
Obs that the FoxD3 data is from normalized data and not counts. Be careful with
the interpretation.

# Read smart-seq data
```{r}
filenameSmrtS <- list.files(paste0(path, "02_Analysis/"), 
           pattern = "02_FilterSmrtS_output_")
if (length(filenameSmrtS) > 1) {
  print(paste0("You have two files matching the pattern.", 
               "Choose one to read and remove the other from the directory."))
  print(filenameSmrtS)
}
```

```{r}
rel_exp.df <- readRDS(paste0(path, "02_Analysis/", filenameSmrtS)) %>%
  subset(orig.ident == "Plate_09") %>%
  FetchData(vars = c("SOX10",
                     "SOX9",
                     "SNAI1",
                     "SNAI2",
                     "TFAP2A",
                     "nCount_RNA"), layer = "counts") %>%
  mutate(Sox10_PM  = SOX10  / (nCount_RNA / 1e+06),
         Sox9_PM   = SOX9   / (nCount_RNA / 1e+06),
         Snai1_PM  = SNAI1  / (nCount_RNA / 1e+06),
         Snai2_PM  = SNAI2  / (nCount_RNA / 1e+06),
         Tfap2a_PM = TFAP2A / (nCount_RNA / 1e+06),
         Dataset = "Strictly sorted",
         Sorted = "HNK-1 top 5%") %>%
  select(c("Sox10_PM", "Sox9_PM", "Snai1_PM", "Snai2_PM", 
           "Tfap2a_PM", "Dataset", "Sorted")) %>%
  rbind(rel_exp.df)

rel_exp.df <- readRDS(paste0(path, "02_Analysis/", filenameSmrtS)) %>%
  subset(orig.ident == "Plate_11") %>%
  FetchData(vars = c("SOX10",
                     "SOX9",
                     "SNAI1",
                     "SNAI2",
                     "TFAP2A",
                     "nCount_RNA"), layer = "counts") %>%
  mutate(Sox10_PM  = SOX10  / (nCount_RNA / 1e+06),
         Sox9_PM   = SOX9   / (nCount_RNA / 1e+06),
         Snai1_PM  = SNAI1  / (nCount_RNA / 1e+06),
         Snai2_PM  = SNAI2  / (nCount_RNA / 1e+06),
         Tfap2a_PM = TFAP2A / (nCount_RNA / 1e+06),
         Dataset = "Strictly sorted",
         Sorted = "HNK-1 top 5%") %>%
  select(c("Sox10_PM", "Sox9_PM", "Snai1_PM", "Snai2_PM", 
           "Tfap2a_PM", "Dataset", "Sorted")) %>%
  rbind(rel_exp.df)
rm(filenameSmrtS); gc()
```

# Plot all data
## Calc summaries for plotting
```{r}
#Compute the some summaries to be used for plotting
summary_stats <- rel_exp.df %>%
  group_by(Dataset, Sorted) %>%
  summarize(#Sample size
            n = n(),
            n_char = paste0("n = ", n()),
            #Sox10
            n_none0_Sox10 = sum(Sox10_PM > 0, na.rm =TRUE),
            perc_non0_Sox10 = (sum(Sox10_PM > 0, na.rm =TRUE)/n())*100,
            perc_non0_Sox10_char = paste0("% non-zero cells = ", 
                                       (sum(Sox10_PM > 0, na.rm =TRUE)/n()) %>%
                                         round(3)*100),
            #Sox9
            n_none0_Sox9 = sum(Sox9_PM > 0, na.rm =TRUE),
            perc_non0_Sox9 = (sum(Sox9_PM > 0, na.rm =TRUE)/n())*100,
            perc_non0_Sox9_char = paste0("% non-zero cells = ", 
                                       (sum(Sox9_PM > 0, na.rm =TRUE)/n()) %>%
                                         round(3)*100),
            #Snai1
            n_none0_Snai1 = sum(Snai1_PM > 0, na.rm =TRUE),
            perc_non0_Snai1 = (sum(Snai1_PM > 0, na.rm =TRUE)/n())*100,
            perc_non0_Snai1_char = paste0("% non-zero cells = ", 
                                       (sum(Snai1_PM > 0, na.rm =TRUE)/n()) %>%
                                         round(3)*100),
            #Snai2
            n_none0_Snai2 = sum(Snai2_PM > 0, na.rm =TRUE),
            perc_non0_Snai2 = (sum(Snai2_PM > 0, na.rm =TRUE)/n())*100,
            perc_non0_Snai2_char = paste0("% non-zero cells = ", 
                                       (sum(Snai2_PM > 0, na.rm =TRUE)/n()) %>%
                                         round(3)*100),
            #TFAP2A
            n_none0_Tfap2a = sum(Tfap2a_PM > 0, na.rm =TRUE),
            perc_non0_Tfap2a = (sum(Tfap2a_PM > 0, na.rm =TRUE)/n())*100,
            perc_non0_Tfap2a_char = paste0("% non-zero cells = ", 
                                       (sum(Tfap2a_PM > 0, na.rm =TRUE)/n()) %>%
                                         round(3)*100)
            )

# Reorder the Dataset factor levels based on perc_non0_Sox10
rel_exp.df$Dataset <- factor(rel_exp.df$Dataset, 
          levels = summary_stats$Dataset[order(summary_stats$perc_non0_Sox10)])
summary_stats
```

## Sox10
```{r}
P1 <- ggplot(rel_exp.df, aes(y = Dataset, x = Sox10_PM, 
                             fill = Sorted)) +
  ggrastr::rasterise(dpi = 1600,
                     layers = c("Point"),
                     input = geom_density_ridges(#bandwidth = .3,
                                    jittered_points = TRUE,
                                    position = position_raincloud(ygap = .02, 
                                                                  height = .2),
                                    point_shape = 21,
                                    point_alpha = .2,
                                    point_size = 1,
                                    alpha = 1,
                                    rel_min_height = 0.001,
                                    scale = .6)) +
  scale_x_continuous(trans = "log1p",
                     limits = c(0, 12050)) +
  # Set colors
  scale_fill_manual(values = c("HNK-1 top 5%"  = "#64EC09",
                               "Sox10"         = "#00bfc4b2",
                               "HNK-1 top 15%" = "#53AC90",
                               "Whole embryo"  = "#BDBDBD")) +
  #Add percentage non-zero cells and sample size to each data set
  geom_text(data = summary_stats,
            aes(x = 100, y = Dataset, label = perc_non0_Sox10_char),
            vjust = -3.5, size = 8*.35) + # *.35 translates pt to mm
  geom_text(data = summary_stats,
            aes(x = 6000, y = Dataset, label = n_char),
            vjust = -3.5, size = 8*.35) + # *.35 translates pt to mm
  
  #Adjust theme 
  theme(#plot.margin = margin(t=1.7, r=1, b=1.7, l=0.4, unit = "mm"),
        #Axes
        axis.text  = element_text(angle = 45, vjust = 1, hjust = 1, size = 8),
        axis.title.y = element_text(size = 11),
        axis.title.x = element_text(size = 11, vjust = 0),
        axis.text.y = element_text(angle = 45),
        axis.ticks.y = element_blank(),
        #Legend
        legend.text  = element_text(size = 8),
        legend.title = element_text(size = 11),
        legend.key.size = unit(3, "mm"),
        legend.position = "none",
        #Background
        plot.background = element_rect(fill = "white", color = "white"),
        panel.background = element_rect(fill = "white", color = "white"),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray"),
        panel.border = element_rect(colour = "black", fill = NA)) +
  labs(x = "Sox10 counts per 1M total counts",
       y = "Frequency of cells")
P1

ggsave(paste0(path, "03_Results/relExpSox10.svg"),
       P1, height = 160, width = 89, units = "mm")
```

Aquivalent plots could be produced for other NCC markers. The DF here contains four more genes. These tell the same story as Sox10 and so are not shown in the paper. By adjusting the code you could plot even more genes. 

Calculated runtime:
```{r}
end_time <- Sys.time()
paste0("Elapsed time: ", end_time - start_time)
sessionInfo()
```