---
title: "plot_FACS.Rmd"
author: "Robin Pranter"
date: "2024-11-04"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to produce plots of flow cytometry data for the methods
paper.

#Setup
```{r}
options(future.globals.maxSize = 120000 * 1024^2)
Sys.setenv(R_MAX_VSIZE = "120Gb")

rm(list=ls()); gc()

getwd()

path <- "~/WorkDirs/PmurNCCPilot_10X_rerun/"

library(tidyverse)
library(flowCore)
library(ggcyto)
library(patchwork)
library(ggridges)
library(scales)
library(car)          # For Anova(type = "III")

```

# Read data
## Read metadata
```{r}
#Read meta data and remove rows with missing file names
meta.df <- read.csv(file = paste0(path, "01_Data/01_FlowCytMetaData.csv"))
meta.df$FileName <- na_if(meta.df$FileName, "")
meta.df <- subset(meta.df, is.na(FileName) == FALSE)
```

## Read FCS-files
```{r}
#List FCS-files in the directory
#filelist <- dir(pattern = ".fcs", full.names = FALSE)
#Check that all filenames are the same in the directory and the metadata
#filelist %in% meta.df$FileName
#meta.df$FileName %in% filelist
#replace the filelist with the FileName column in the metadata
filelist <- paste0(path, "01_Data/", meta.df$FileName)

#Check that all the listed files are actually FCS-files.
if (isFCSfile(files = filelist) %>% sum() == length(filelist)){
  ##If they are FCS-files: Read all FCS-files to list
  data.list <- lapply(filelist, read.FCS, 
                      transformation = FALSE, alter.names = TRUE)
  names(data.list) <- meta.df$FileName
  ##Some of the fcs-files have two more data columns "DAPI.W" and "APC"(Sox10)
  ##All flowFrames in a flowSet need to have the same number of variables
  ###Loop over data.list
  for (i in 1:length(data.list)){
    ###If florFrame has less than 12 variables, add "DAPI.W" and "APC"
    if (data.list[[i]] %>% exprs() %>% ncol() < 12){
      new.cols <- matrix(rep(NA, times = 2*nrow(data.list[[i]])),
                         ncol = 2)
      colnames(new.cols) <- c("DAPI.W", "APC.A")
      data.list[[i]] <- fr_append_cols(fr = data.list[[i]], new.cols)
    }
    ##Reorder the columns so all flowFrames have the columns in the same order
    exprs(data.list[[i]]) <- exprs(data.list[[i]])[,c("FSC.A", "FSC.H", "FSC.W", 
                                                      "SSC.A", "SSC.H", "SSC.W", 
                                                      "FITC.A", "APC.Cy7.A", 
                                                      "DAPI.A", "DAPI.W", 
                                                      "APC.A", "Time")]
  }
  ##Create a flowSet from data.list
  data.fs <- as(data.list, "flowSet")
} else {
  #Otherwise: print error message
  print(paste0("One of the files: ", filelist, " is not an FCS-file"))
}
rm(i, new.cols); gc()
```

## Subset data
```{r}
# This is based on manually going through all the scatter plots and lab 
# notebooks to see which data is reliable
#plot.fs <- data.fs[c(6:56, 62:82)]
#data.fs <- data.fs[c(6:9, 11:19, 22, 24:29, 34:52, 62:66, 79:80)]

# Excluding those with no LD-stain or different incubation times
data.fs <- data.fs[c(6:9, 11:19, 22, 24:29, 34:52, 65:66, 79:80)]
gc()
```

# Remove debri 
## Construct 1st gate
```{r}
#Polygon gate (approx like by the instrument)
#g1 <- matrix(c(30, 40, 120, 210, 200, 90, 40,
#               10,  5,  10,  30,  80, 70, 20), 
#             ncol = 2, nrow = 7) * 1000
#colnames(g1) <- c("FSC.A","SSC.A")
#g1
#g1 <- polygonGate(filterId = "P1", boundaries = g1)
#g1
#data.ff_g1 <- flowCore::Subset(data.ff, g1)

#Ellipsoid gate
#exprs(data.ff)[, c("FSC.A", "SSC.A")] %>% var()
g1_cov <- matrix(c(1.0*10^10, 1.8*10^9,
                   1.8*10^ 9, 1.3*10^9),
                 ncol = 2, nrow = 2,
                 dimnames = list(c("FSC.A","SSC.A"), c("FSC.A","SSC.A")))
#exprs(data.ff)[, "FSC.A"] %>% mean()
#exprs(data.ff)[, "SSC.A"] %>% mean()
g1_mean <- c("FSC.A"=1.2*10^5,"SSC.A"=3.7*10^4)
g1 <- ellipsoidGate(filterId = "g1", .gate = g1_cov, mean = g1_mean)
rm(g1_cov, g1_mean)
data.fs_g1 <- flowCore::Subset(data.fs, g1)
```

### Plot 1st gate
```{r}
p1 <- ggcyto(data.fs, aes(x = FSC.A, y = SSC.A)) + 
  geom_hex(bins = 40) +
  geom_gate(g1) + 
  theme(strip.text = element_text(size = 8),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 11)) + 
  xlab(label = "FSC-A") + 
  ylab(label = "SSC-A")
p1
```


```{r}
p1b <- ggcyto(data.fs@frames$Specimen_001_Pool6.fcs, 
              aes(x = FSC.A, y = SSC.A)) + 
  geom_hex(bins = 40) +
  geom_gate(g1) + 
  #geom_stats(type = c("percent"), location = "plot") + 
  ggtitle(label = paste0("All data (n = ",
                         exprs(data.fs@frames$Specimen_001_Pool6.fcs) %>% 
                           nrow(),
                         ")")) +
  xlab(label = "FSC-A") + 
  ylab(label = "SSC-A") +
  theme_classic() +
  theme(strip.text = element_blank(),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 8),
        panel.background = element_rect(colour = "black", linewidth = .5),
        axis.line = element_blank(),
        legend.position = "none")

#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel1.svg"), 
#       plot = p1b, height = 45, width = 45, units = "mm")

p1b
```


## Construct 2nd gate
```{r}
#Polygon gate (approx like by the instrument)
#g2 <- matrix(c(10, 100, 120, 120, 110,  70,  40,  20,
#               60,  90, 100, 120, 130, 130, 110, 100), 
#             ncol = 2, nrow = 8) * 1000
#colnames(g2) <- c("FSC.H","FSC.W")
#g2
#g2 <- polygonGate(filterId = "g2", boundaries = g2)
#g2
#data.ff_g2 <- flowCore::Subset(data.ff_g1, g2)

#Ellipsoid gate
#exprs(data.ff_g1)[, c("FSC.H","FSC.W")] %>% var()
g2_cov <- matrix(c(3*10^9, 4*10^8,
                   4*10^8, 2*10^9),
                 ncol = 2, nrow = 2,
                 dimnames = list(c("FSC.H","FSC.W"), c("FSC.H","FSC.W")))
#exprs(data.ff_g1)[, "FSC.H"] %>% mean()
#exprs(data.ff_g1)[, "FSC.W"] %>% mean()
g2_mean <- c("FSC.H"=6.4*10^4,"FSC.W"=8.9*10^4)
g2 <- ellipsoidGate(filterId = "Gate 2", .gate = g2_cov, mean = g2_mean)
rm(g2_cov, g2_mean)
data.fs_g2 <- flowCore::Subset(data.fs_g1, g2)
```

## Remove doublets and debri
### Plot 2nd gate
```{r}
p2 <- ggcyto(data.fs_g1, aes(x = FSC.H, y = FSC.W)) + 
  geom_hex(bins = 40) +
  geom_gate(g2) + 
  theme(axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 11)) +
  ggtitle(label = "Gate 1") +
  xlab(label = "FSC-H") + 
  ylab(label = "FSC-W")
p2
```

```{r}
p2b <- ggcyto(data.fs_g1@frames$Specimen_001_Pool6.fcs, 
              aes(x = FSC.H, y = FSC.W)) + 
  geom_hex(bins = 40) +
  geom_gate(g2) +
  #geom_stats(type = c("percent"), location = "plot") + 
  ggtitle(label = paste0("Gate 1 (n = ",
                         exprs(data.fs_g1@frames$Specimen_001_Pool6.fcs) %>% 
                           nrow(),
                         ")")) +
  xlab(label = "FSC-H") + 
  ylab(label = "FSC-W") +
  theme_classic() +
  theme(strip.text = element_blank(),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 8),
        panel.background = element_rect(colour = "black", linewidth = .5),
        axis.line = element_blank(),
        legend.position = "none")

#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel2.png"),
#       plot = p2b, height = 45, width = 45, units = "mm")  

p2b
```

### Plot 2nd gate results
```{r}
p3 <- ggcyto(data.fs_g2, aes(x = FSC.A, y = FSC.H)) + 
  geom_hex(bins = 40)+ 
  #theme(strip.text = element_blank()) + 
  theme(axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 11)) +
  ggtitle(label = "Gate 2") +
  xlab(label = "FSC-A") + 
  ylab(label = "FSC-H")
p3
```

```{r}
p3b <- ggcyto(data.fs_g2@frames$Specimen_001_Pool6.fcs, 
              aes(x = FSC.A, y = FSC.H)) + 
  geom_hex(bins = 40)+ 
  ggtitle(label = paste0("Gate 2 (n = ",
                         exprs(data.fs_g2@frames$Specimen_001_Pool6.fcs) %>% 
                           nrow(), 
                         ")")) +
  xlab(label = "FSC-A") + 
  ylab(label = "FSC-H") +
  theme_classic() +
  theme(strip.text = element_blank(),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 8),
        panel.background = element_rect(colour = "black", linewidth = .5),
        axis.line = element_blank(),
        legend.position = "none")

#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel3.svg"),
#       plot = p3b, height = 45, width = 45, units = "mm")

p3b
```


# Remove dead cells
## Construct 3rd gate
```{r}
g3 <- matrix(c(3.5*10^4, 2.45*10^5,
                 -500,      550), 
             ncol = 2, nrow = 2)
colnames(g3) <- c("DAPI.A","APC.Cy7.A")
g3
g3 <- rectangleGate(filterId = "g3", boundaries = g3)
g3
data.fs_g3 <- flowCore::Subset(data.fs_g2, g3)
```

### Plot 3rd gate
#### Subset plot data
```{r}
## Only keep data sets with >= 15000 cells
x <-c()
for (i in 1:length(data.fs_g2)){x[i] <- data.fs_g2[[i]] %>% exprs() %>% nrow()}
x <- x >= 15000
data.fs_g2 <- data.fs_g2[x]
rm(i,x)
```

#### Plot
```{r}
p4 <- ggcyto(data.fs_g2, aes(x = DAPI.A, y = APC.Cy7.A)) + 
  geom_hex(bins = 40) + 
  geom_gate(g3) + 
  #xlim(0, 10^6)+
  scale_x_logicle() +
  scale_y_logicle()+ 
  #theme(strip.text = element_blank()) + 
  theme(axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 11)) +
  ggtitle(label = "Gate 2") +
  xlab(label = "DAPI") + 
  ylab(label = "Live/Dead (APC-Cy7)")
p4
```

```{r}
p4b <- ggcyto(data.fs_g2@frames$Specimen_001_Pool6.fcs, 
              aes(x = DAPI.A, y = APC.Cy7.A)) + 
  geom_hex(bins = 40) + 
  #geom_stats(type = c("percent", "count"), location = "plot") +
  geom_gate(g3) + 
  #xlim(0, 10^6)+
  scale_x_logicle() +
  scale_y_logicle()+ 
  ggtitle(label = paste0("Gate 2 (n = ",
                         exprs(data.fs_g2@frames$Specimen_001_Pool6.fcs) %>% 
                           nrow(),
                         ")")) +
  xlab(label = "DAPI") + 
  ylab(label = "Live/Dead") +
  theme_classic() +
  theme(strip.text = element_blank(),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 8),
        panel.background = element_rect(colour = "black", linewidth = .5),
        axis.line = element_blank(),
        legend.position = "none")

#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel4.svg"),
#       plot = p4b, height = 45, width = 45, units = "mm")
  
p4b
```

### Plot 3rd gate results
```{r}
data.fs_g3 <- data.fs_g3[!data.fs_g3@frames %>% names() %in% 
                        c("Specimen_001_#5_1-800.fcs", 
                          "Specimen_001_#1_1-50.fcs", 
                          "Specimen_001_#4_1-400.fcs", 
                          "Specimen_001_#3_1-200.fcs")]
p6 <- ggcyto(data.fs_g3, aes(x = FITC.A, y = FSC.A)) + 
  geom_hex(bins = 40) +
  #geom_hex(binwidth = c(.02, 1000)) + #binwidth = c(vertical, horizontal)
  scale_x_logicle() + 
  #theme(strip.text = element_blank()) + 
  theme(axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1)) +
  theme(axis.title = element_text(size = 11)) +
  theme(plot.title = element_text(size = 11)) +
  ggtitle(label = "Gate 3") +
  xlab(label = "HNK-1 (FITC)") + 
  ylab(label = "FSC-A")
p6
```

```{r}
p6b <- ggcyto(data.fs_g3@frames$Specimen_001_Pool6.fcs, 
              aes(x = FITC.A, y = FSC.A)) + 
  geom_hex(bins = 40) +
  scale_x_logicle() + 
  ggtitle(label = paste0("Gate 3 (n = ",
                         exprs(data.fs_g3@frames$Specimen_001_Pool6.fcs) %>% 
                           nrow(),
                         ")")) +
  xlab(label = "HNK-1 (FITC)") + 
  ylab(label = "FSC-A") +
  theme_classic() +
  theme(strip.text = element_blank(),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 8),
        panel.background = element_rect(colour = "black", linewidth = .5),
        axis.line = element_blank(),
        legend.position = "none")

#ggsave(filename = paste0(path, "../02_FlowCytResults/","FlowPlot_Panel6.svg"), 
#       plot = p6b, height = 45, width = 90, units = "mm")
  
p6b
```

```{r}
p6c <- ggcyto(data.fs_g3@frames$Specimen_001_DAPI_control.fcs , 
              aes(x = FITC.A, y = FSC.A)) + 
  geom_hex(bins = 20) +
  scale_x_logicle() + 
  coord_cartesian(xlim = c(-1e2, 1e4), ylim = c(0, 200000)) +
  ggtitle(label = paste0("Gate 3 (n = ",
                         exprs(data.fs_g3@frames$Specimen_001_Pool6.fcs) %>% 
                           nrow(),
                         ")")) +
  xlab(label = "HNK-1 (FITC)") + 
  ylab(label = "FSC-A") +
  theme_classic() +
  theme_classic() +
  theme(strip.text = element_blank(),
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.title = element_text(size = 11),
        plot.title = element_text(size = 8),
        panel.background = element_rect(colour = "black", linewidth = .5),
        axis.line = element_blank(),
        legend.position = "none")

#ggsave(filename = paste0(path, "../02_FlowCytResults/","FlowPlot_Panel6c.svg"), 
#       plot = p6c, height = 45, width = 90, units = "mm")
  
p6c
```

### Plot DAPI and FITC histogram
#### Subset plot data
```{r}
## Only keep data sets with >= 15000 cells
x <-c()
for (i in 1:length(data.fs_g3)){x[i] <- data.fs_g3[[i]] %>% exprs() %>% nrow()}
x <- x >= 15000
data.fs_g3 <- data.fs_g3[x]
rm(i, x)
```

##### DAPI df
```{r}
# Extract the "DAPI.A" channel from each flowFrame
dapi_values <- lapply(seq_along(data.fs_g3), function(i) {
  # Extract data for the "DAPI.A" channel from the i-th flowFrame
  flowFrame_data <- exprs(data.fs_g3[[i]])
  return(flowFrame_data[, "DAPI.A"])
})

# Combine all the DAPI values into a single vector and create a corresponding 
# label
dapi_combined <- unlist(dapi_values)
flowFrame_labels <- rep(1:length(data.fs_g3), sapply(dapi_values, length))
flowFrame_labels <- names(data.fs_g3@frames)[flowFrame_labels]

# Create a data frame
dapi.df <- data.frame(DAPI_A = dapi_combined, FlowFrame = factor(flowFrame_labels))
rm(dapi_values, flowFrame_labels, dapi_combined)

# Add somite stage
dapi.df <- merge(dapi.df, meta.df[,c("FileName", "SomStage")], 
                 by.x = "FlowFrame", by.y = "FileName", all.x = TRUE)

# Sort dapi.df by the SomStage column (assuming SomStage is numeric)
dapi.df <- dapi.df %>%
  arrange(SomStage) %>%  # Sort by SomStage
  # Reorder the FlowFrame factor levels
  mutate(FlowFrame = factor(FlowFrame, levels = unique(FlowFrame)))  

```

##### FITC df
```{r}
# Extract the "FITC.A" channel from each flowFrame
fitc_values <- lapply(seq_along(data.fs_g3), function(i) {
  # Extract data for the "FITC.A" channel from the i-th flowFrame
  flowFrame_data <- exprs(data.fs_g3[[i]])
  return(flowFrame_data[, "FITC.A"])
})

# Combine all the FITC and values into a single vector and create a 
# corresponding label
fitc_combined <- unlist(fitc_values)
flowFrame_labels <- rep(1:length(data.fs_g3), sapply(fitc_values, length))
flowFrame_labels <- names(data.fs_g3@frames)[flowFrame_labels]

# Create a data frame
fitc.df <- data.frame(FITC_A = fitc_combined, 
                      FlowFrame = factor(flowFrame_labels))

# Add somite stage
fitc.df <- merge(fitc.df, meta.df[,c("FileName", "SomStage", "Phen")], 
                 by.x = "FlowFrame", by.y = "FileName", all.x = TRUE)

# Sort fitc.df by the SomStage column
fitc.df <- fitc.df %>%
  arrange(SomStage) %>%  # Sort by SomStage
  # Reorder factor levels
  mutate(FlowFrame = factor(FlowFrame, levels = unique(FlowFrame))) 
```

```{r}
dapi.df$FlowFrame %>% unique() %>% length()
fitc.df$FlowFrame %>% unique() %>% length()
```

#### Plot DAPI
```{r}
# Create the ridgeplot using ggplot2 and ggridges
p5  <- ggplot(dapi.df, aes(x = DAPI_A, y = FlowFrame, fill = SomStage)) + 
  geom_density_ridges(scale = 6, alpha = .7) +
  scale_x_logicle(breaks = c(3*10^4, 6*10^4, 9*10^4, 1.2*10^5, 1.5*10^5, 
                             1.8*10^5, 2.1*10^5, 2.4*10^5)) +
  labs(title = "Gate 3", x = "DAPI", y = "Density", fill = "Somites") +
  #scale_y_discrete(labels = c("qPCR", "qPCR", rep("scRNA-seq", 6))) +
  scale_y_discrete(labels = c(rep("Short label", 
                                  length(unique(dapi.df$FlowFrame))))) +
  guides(fill = guide_legend(reverse = FALSE, nrow = 1)) +
  #Change some themes
  theme(plot.margin = margin(t=1.7, r=2, b=1.7, l=0.4, unit = "mm"),
        #Axes
        axis.text = element_text(angle = 45, size = 5, vjust = 1, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        #Titles
        axis.title.x = element_text(hjust = 0.5, size = 8),
        axis.title.y = element_text(hjust = 0.5, size = 8),
        plot.title = element_text(size = 8, face = "plain"),
        #legend.position = "none",
        # Legend
        legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 5),
        legend.key.size = unit(2.5, "mm"),
        
        #Background
        plot.background = element_rect(fill = "white", color = "white"),
        panel.background = element_rect(fill = "white", color = "white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA)
        )
p5

#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel5.svg"), 
#       plot = p5, height = 112.7083, width = 56.750, units = "mm")
```

#### Plot FITC
```{r}
p7  <- ggplot(subset(fitc.df, FITC_A > -50 & FITC_A < 3*10^4), 
              aes(x = FITC_A, y = FlowFrame, fill = SomStage)) + 
  geom_density_ridges(scale = 5, alpha = .7) +
  scale_x_logicle(breaks = c(0, 10^2, 10^3, 10^4)) +
  labs(title = "Gate 3", x = "HNK-1 (FITC)", y = "Density", fill = "Somites") +
  #scale_y_discrete(labels = c(rep("Short label", 
  #                                length(unique(dapi.df$FlowFrame))))) +
  guides(fill = guide_legend(reverse = FALSE, nrow = 1)) +
  #Change some themes
  theme(plot.margin = margin(t=1.7, r=1, b=1.7, l=0.4, unit = "mm"),
        # Axes
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        # Titles
        axis.title.x = element_text(hjust = 0.5, size = 11),
        axis.title.y = element_text(hjust = 0.5, size = 11),
        plot.title = element_text(size = 11, face = "plain"),
        # Legend
        legend.position = "bottom",
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 8),
        legend.key.size = unit(3, "mm"),
        
        #Background
        plot.background = element_rect(fill = "white", color = "white"),
        panel.background = element_rect(fill = "white", color = "white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA)
        )
p7
#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel7.svg"),
#      plot = p7, height = 112.7083, width = 115.08335, units = "mm")
```

##### Again, but centered
Center the peak of each ridge at 0
```{r}
#fitc.df <- fitc.df %>% select(-peak, -FITC_logicle, -FITC_centered_logicle)

# Apply logicle transform before centering
logicle_transform <- flowCore::logicleTransform(w = 0.5, t = 262144, 
                                                m = 4.5, a = 0)
fitc.df <- fitc.df %>%
  mutate(FITC_logicle = logicle_transform(FITC_A))

# Compute peak on transformed values
peaks <- fitc.df %>%
  group_by(FlowFrame) %>%
  summarize(peak = {
    d <- density(FITC_logicle, n = 2^7, bw = 0.0291)#0.0557)
    d$x[which.max(d$y)]
  })

# Center logicle-transformed values
fitc.df <- fitc.df %>%
  left_join(peaks, by = "FlowFrame") %>%
  mutate(FITC_centered_logicle = FITC_logicle - peak)
```

```{r}
p8 <- ggplot(fitc.df,
             aes(x = FITC_centered_logicle, 
                 y = FlowFrame, 
                 fill = SomStage)) + 
  # Draw ridges
  geom_density_ridges(scale = 5, alpha = .7) +
  
  # Set breaks and x-limits to mimic the plots above
  scale_x_continuous(
    breaks = logicle_transform(c(0, 1e2, 1e3, 1e4)) - logicle_transform(0),
    labels = c("0", "10²", "10³", "10⁴")) +
  coord_cartesian(xlim=logicle_transform(c(-100, 1e4)) - logicle_transform(0)) +
  #geom_vline(xintercept = 0) +

  #Set aesthetics
  labs(title = "Gate 3", x = "HNK-1 (FITC)", y = "Density", fill = "Somites") +
  guides(fill = guide_legend(reverse = FALSE, nrow = 1)) +
  theme(plot.margin = margin(t=1.7, r=1, b=1.7, l=0.4, unit = "mm"),
        # Axes
        axis.text = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        # Titles
        axis.title.x = element_text(hjust = 0.5, size = 11),
        axis.title.y = element_text(hjust = 0.5, size = 11),
        plot.title = element_text(size = 8, face = "plain"),
        # Legend
        legend.position = "none",
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 8),
        legend.key.size = unit(3, "mm"),
        #Background
        plot.background = element_rect(fill = "white", color = "white"),
        panel.background = element_rect(fill = "white", color = "white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))

#ggsave(filename = paste0(path, "../02_FlowCytResults/", "FlowPlot_Panel8.svg"),
#      plot = p8, height = 80, width = 90, units = "mm")

p8
```

# qPCR
## Construct sort gates
```{r}
g_nega <- matrix(c( -120,    248,
                   20000, 194000),
                 ncol = 2, nrow = 2)
colnames(g_nega) <- c("FITC.A","FSC.A")
g_nega <- rectangleGate(filterId = "g_nega", boundaries = g_nega)

g_10pe <- matrix(c(  248,    477,
                   20000, 194000),
                 ncol = 2, nrow = 2)
colnames(g_10pe) <- c("FITC.A","FSC.A")
g_10pe <- rectangleGate(filterId = "g_10pe", boundaries = g_10pe)

g_5per <- matrix(c(  477,    880,
                   20000, 194000),
                 ncol = 2, nrow = 2)
colnames(g_5per) <- c("FITC.A","FSC.A")
g_5per <- rectangleGate(filterId = "g_5per", boundaries = g_5per)

g_2.5p <- matrix(c(  880,  14500,
                   20000, 194000),
                 ncol = 2, nrow = 2)
colnames(g_2.5p) <- c("FITC.A","FSC.A")
g_2.5p <- rectangleGate(filterId = "g_2.5p", boundaries = g_2.5p)
```

## Plot sorting gates
```{r}
p9 <- ggcyto(data.fs_g3@frames$Specimen_001_Pool6.fcs, 
             aes(x = FITC.A, y = FSC.A)) + 
  geom_hex(bins = 40) +
  scale_x_logicle() +
  geom_gate(g_nega) + 
  geom_gate(g_10pe) +
  geom_gate(g_5per) +
  geom_gate(g_2.5p) +
  geom_stats(type = "percent", digits = 2) + 
  theme(strip.text = element_blank()) + 
  theme(axis.text = element_text(angle = 45, size = 5, vjust = 1, hjust = 1)) +
  theme(axis.title = element_text(size = 8)) +
  theme(plot.title = element_text(size = 8)) +
  ggtitle(label = "Gate 3") +
  xlab(label = "HNK-1 (FITC)") + 
  ylab(label = "FSC-A")
p9
```

