---
title: ''
output: html_document
date: "2025-12-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up
```{r}
options(future.globals.maxSize = 120000 * 1024^2) # Change this to fit your 
Sys.setenv(R_MAX_VSIZE = "120Gb")                 # computer

rm(list=ls());gc()
start_time <- Sys.time()

library(Seurat, lib.loc = "/usr/local/lib/R/site-library")    #Change these to
library(glmGamPoi, lib.loc = "/usr/local/lib/R/site-library") #mach the location
library(dplyr, lib.loc = "/usr/local/lib/R/site-library")     #of your libraries
library(ggplot2, lib.loc = "/usr/local/lib/R/site-library") 
library(patchwork, lib.loc = "/usr/local/lib/R/site-library")
library(latex2exp, lib.loc = "/home/pranter/R/x86_64-pc-linux-gnu-library/4.5")
library(ggrepel, lib.loc = "/usr/local/lib/R/site-library")
library(ggbreak, lib.loc = "/home/pranter/R/x86_64-pc-linux-gnu-library/4.5")
library(ggridges, lib.loc = "/usr/local/lib/R/site-library")
library(stringr, lib.loc = "/usr/local/lib/R/site-library")

path <- "~/WorkDirs/PmurNCCPilot_10X_rerun/"
```

# Read filtered 10X data
```{r}
filename <- list.files(paste0(path, "02_Analysis/"), 
           pattern = "03_CompileAndFilter10X_out_")
if (length(filename) == 1) {
  TenX.SO <- readRDS(file=paste0(path, "02_Analysis/", filename))
} else {
  print(paste0("You have two files matching the pattern.", 
               "Choose one to read and remove the other from the directory."))
  print(filename)
}
rm(filename)
TenX.SO
```

# Read filtered SmartSeq data
```{r}
filename <- list.files(paste0(path,"02_Analysis/"),
           pattern = "02_FilterSmrtS_output_")
if (length(filename) == 1) {
  SmrtS.SO <- readRDS(file = paste0(path, "02_Analysis/", filename))
} else {
  print(paste0("You have two files matching the pattern. ", 
               "Choose one to read and remove the other from the directory."))
  print(filename)
}
rm(filename)
SmrtS.SO
```

# Merge objects
```{r}
# Select metadata columns
met.cols <- c("orig.ident", "type", "nCount_RNA","nFeature_RNA", 
              "percent.mt", "percent.rb")
SmrtS.SO$type <- "Strictly sorted"
SmrtS.SO@meta.data <- SmrtS.SO@meta.data %>% select(all_of(met.cols))
TenX.SO@meta.data <- TenX.SO@meta.data %>% select(all_of(met.cols))
rm(met.cols)

#Restor seurat objects to non-normalized form
DefaultAssay(SmrtS.SO) <- "RNA"
SmrtS.SO <- DietSeurat(object = SmrtS.SO,
                layers = "counts",
                assays = "RNA")
DefaultAssay(TenX.SO) <- "RNA"
TenX.SO <- DietSeurat(object = TenX.SO,
                layers = "counts",
                assays = "RNA")

#Merge
full.SO <- merge(TenX.SO, c(SmrtS.SO))
full.SO[["RNA"]] <- JoinLayers(full.SO[["RNA"]])

# Fix the type column in meta data (mostly for plotting)
full.SO@meta.data$type <- factor(full.SO@meta.data$type, 
                                 levels = c("Unsort", "Sorted", 
                                            "Strictly sorted"))
full.SO@meta.data <- full.SO@meta.data %>%
  mutate(type = recode(type, 
                       "Unsort" = "Unsorted",
                       "Sorted" = "Leniently sorted"),
         type.short = recode(type, 
                             "Unsorted" = "U",
                             "Leniently sorted" = "L",
                             "Strictly sorted" = "S"))

full.SO
full.SO@meta.data %>% colnames()
```

# Basic stats
```{r}
full.SO@meta.data %>% 
  group_by(type) %>% 
  summarize(n = n(),
            av_count = mean(nCount_RNA) %>% round(),
            sd_count = sd(nCount_RNA) %>% round(),
            av_genes = mean(nFeature_RNA) %>% round(),
            sd_genes = sd(nFeature_RNA) %>% round()) %>%
  t()
```

# Calc Cell Cycl
```{r}
# temporary normalisation for CC-scoring
full.SO <- SCTransform(full.SO,
                       assay = 'RNA',
                       new.assay.name = 'temp_SCT',
                       vars.to.regress = c('percent.mt'),
                       vst.flavor = "v2",
                       variable.features.n = 3000) #3000 is default
# Load ortholog table
Ortho.df <- read.csv(file=paste0(path, "01_Data/", 
                            "Pmur1.0Human_OrthoTable_BioMartManDL20241206.txt"))
# Save S and G2M markers in Pmur
G2M.genes <- Ortho.df[Ortho.df$Human.gene.name %in% 
                        cc.genes.updated.2019$g2m.genes, ]$Gene.name
S.genes <- Ortho.df[Ortho.df$Human.gene.name %in% 
                      cc.genes.updated.2019$s.genes, ]$Gene.name

#Check that the markers are present in the data
print(paste0(full.SO %>% Features() %in% G2M.genes %>% sum(), "/",
             G2M.genes %>% length(), 
             " G2M-markers are present in the data"))
print(paste0(full.SO %>% Features() %in% S.genes %>% sum(), "/",
             S.genes %>% length(), 
             " S-markers are present in the data"))

#Subset only the genes that are present
G2M.genes <- G2M.genes[G2M.genes %in% Features(full.SO)]
S.genes <- S.genes[S.genes %in% Features(full.SO)]
# Predict cell cycle state (make sure to specify the "assay" parameter)
full.SO <- CellCycleScoring(full.SO,
                            s.features = S.genes,
                            g2m.features = G2M.genes,
                            assay = 'temp_SCT',
                            set.ident = TRUE)
rm(G2M.genes, S.genes); gc()
```

# Plot QCs
```{r}
PQC <- c("nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rb") %>% 
  lapply(function(feats){
  ggplot(full.SO@meta.data, aes(x = type, y = .data[[feats]], fill = type)) + 
      geom_violin() +
      scale_fill_manual(values = c("Unsorted"          = "#BDBDBD",
                                   "Leniently sorted"  = "#53AC90",
                                   "Strictly sorted"   = "#64EC09")) +
      labs(y = c("nCount_RNA"   = "n counts",
                 "nFeature_RNA" = "n genes",
                 "percent.mt"   = "% mitochondrial",
                 "percent.rb"   = "% ribosomal")[feats], fill = "") +
      theme_classic() + 
      theme(plot.margin = margin(t=1.7, r=1, b=0, l=0.4, unit = "mm"),
          axis.text.y = element_text(size = 8, angle = 45),
          axis.title.y = element_text(size = 11),
          
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          
          legend.text = element_text(size = 8),
          legend.key.size = unit(3, "mm"),
          legend.title = element_text(size = 11),
          
          plot.title = element_blank())
  }) %>% 
  wrap_plots(ncol = 4) + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
PQC
ggsave(filename = paste0(path, "03_Results/QCPlot.svg"),
      plot = PQC, height = 50, width = 180, units = "mm")
```

# Table for manuscript
```{r}
full.SO@meta.data %>% select("type", "nCount_RNA", "nFeature_RNA") %>% 
  group_by(type) %>% 
  summarize(n = n(),
            av_count = mean(nCount_RNA) %>% round(),
            sd_count = sd(nCount_RNA) %>% round(),
            av_genes = mean(nFeature_RNA) %>% round(),
            sd_genes = sd(nFeature_RNA) %>% round()) %>%
  t()
```

# Standard processing
## Without integration
```{r}
#DefaultAssay(full.SO) <- "RNA"
#full.SO[["temp_SCT"]] <- NULL
## Normalize
#full.SO <- SCTransform(full.SO, 
#                       assay = 'RNA',
#                       new.assay.name = 'SCT',
#                       vars.to.regress =c("percent.mt", "S.Score", "G2M.Score"),
#                       vst.flavor = "v2",
#                       variable.features.n = 3000)
#
## Inspect the normalization
#cbind(Matrix::rowSums(full.SO[["RNA"]]$counts)[
#                                        rownames(full.SO[["RNA"]]$counts) %in% 
#                                          rownames(full.SO[["SCT"]]$data)],
#             Matrix::rowSums(full.SO[["SCT"]]$data)) %>%
#  as.data.frame() %>%
#  dplyr::rename(raw = V1, norm = V2) %>%
#  ggplot(aes(raw, norm)) +
#  geom_point() +
#  labs(title = "Normalized data vs raw counts (each dot is a gene)") +
#  theme(text = element_text(size = 6))
#
## Inspect HVG selection
#VariableFeaturePlot(full.SO, assay = "SCT") %>% 
#  LabelPoints(points = VariableFeatures(full.SO, assay = "SCT") %>% head(20), 
#              repel = TRUE,
#              xnudge = 0, ynudge = 0, max.overlaps = 22) +
#  theme(text = element_text(size = 6))
#
## Lin dim reduc
#full.SO <- RunPCA(full.SO, assay = "SCT")
#ElbowPlot(full.SO, reduction = "pca", ndims = 50)
#
#Non-lin dim reduc
#full.SO <- RunUMAP(full.SO, dims = 1:20)#, spread = 0.4, min.dist = 0.4)
#  # dims = 11 seems good judgeing by the elbowplot and PC-heatmaps 
#  # Changing the min.dist dramatically changes the umap. This requires some 
#  # optimaization. Come back to this later.
#
##Plot
#DimPlot(full.SO, reduction = "umap", group.by = "type", 
#        shuffle = TRUE, alpha = 0.5) + ggtitle("No integration")
##DimPlot(full.SO, group.by = "orig.ident", shuffle = TRUE, alpha = 0.5,
##        cells = Cells(subset(full.SO, subset = type.short == "L")))
```

## With integration
See insturctions here: https://github.com/satijalab/seurat/issues/7542
and here: https://satijalab.org/seurat/articles/seurat5_integration
```{r}
# Normalize each sample separately
full.SO[["RNA"]] <- split(full.SO[["RNA"]], f = full.SO$orig.ident)
DefaultAssay(full.SO) <- "RNA"
full.SO <- SCTransform(full.SO,
                       assay = "RNA",
                       new.assay.name = "SCT_BySample",
                       vars.to.regress =c("percent.mt", "G2M.Score", "S.Score"),
                       vst.flavor = "v2",
                       variable.features.n = 3000) %>%
  # Calculate PCA
  RunPCA(assay = "SCT_BySample", reduction.name = "pca_SCT_BySample") %>%
  # Integrate data sets
  IntegrateLayers(method = HarmonyIntegration,
                  orig.reduction = "pca_SCT_BySample", 
                  new.reduction = 'harmony',
                  assay = "SCT_BySample",
                  normalization.method = "SCT",
                  verbose = FALSE) %>%
  # Calc umap
  RunUMAP(dims = 1:30, reduction = "harmony", reduction.name = "umap_harmony")
```
# Plot sort types in umap
```{r}
P01 <- DimPlot(full.SO, 
               reduction = "umap_harmony", group.by = "type",
               order = c("Strictly sorted", "Leniently sorted", "Unsorted"),
               alpha = 0.5, stroke.size = 0, pt.size = 1) + 
  scale_color_manual(values = c("Unsorted"          = "#BDBDBD",
                                "Leniently sorted"  = "#53AC90",
                                "Strictly sorted"   = "#64EC09")) + 
  labs(y = "umap 2", x = "umap 1") +
  theme(plot.margin = margin(t=1.7, r=1, b=1.7, l=0.4, unit = "mm"),
        axis.text.y = element_text(size = 8, angle = 45),
        axis.title.y = element_text(size = 11),
          
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title.x = element_text(size = 11),

        legend.text = element_text(size = 8),
        legend.key.size = unit(2.5, "mm"),
        legend.position = "inside",
        
        plot.title = element_blank())
P01

ggsave(paste0(path, "03_Results/UMAP_SortType.svg"),
       P01, height = 89, width = 89, units = "mm")
```

# Plot marker exp in umap
```{r}
# First normalize across samples to correct for sequencing depth
full.SO[["RNA"]] <- JoinLayers(full.SO[["RNA"]])
full.SO <- SCTransform(full.SO,
                       assay = "RNA",
                       new.assay.name = "SCT",
                       vars.to.regress =c("percent.mt", "G2M.Score", "S.Score"),
                       vst.flavor = "v2",
                       variable.features.n = 3000)

# It also seems to work to do PrepSCTFindMarkers
```

```{r}
# Then plor using FeaturePlot
P02 <- FeaturePlot(full.SO, c("SOX10", "TFAP2A", 
                              "SNAI2", "FOXD3"#, 
                              #"PAX7", "SOX9", "TWIST1", "PAX3", "GAPDH"
                              ),
            order = TRUE, pt.size = 1, alpha = 0.5, keep.scale = "all") & 
  NoAxes() &
  theme(legend.key.size = unit(3, "mm"),
        legend.text = element_text(size = 8))

#Remove stroke arround points
P02 <- lapply(P02, function(p) {
  p$layers[[1]]$aes_params$stroke <- 0
  old_title <- p$labels$title
#Chance title font
  p <- p + ggtitle(str_to_sentence(old_title)) +
    theme(plot.title = element_text(face = "italic"))
  p
})

#Tidy up
P02 <- wrap_plots(P02, ncol = 2, guides = "collect") & 
  FontSize(main = 11) + 
  NoAxes()

P02

ggsave(paste0(path, "03_Results/UMAP_MrkrExp.svg"),
       P02, height = 89, width = 89, units = "mm")
```
# Calc NCC-ness score
```{r}
full.SO <- AddModuleScore(object = full.SO, 
                          features = list(c("PAX3", "PAX7", "FOXD3", "SOX10",
                                            "SOX9", "WNT1", "TFAP2A", "SOX5",
                                            "SNAI2", "ZIC1")), 
                          name = "NCC_AMS",
                          ctrl = 100)
full.SO@meta.data %>% 
  select(type, NCC_AMS1) %>% 
  group_by(type) %>% 
  summarize(avg_NCCAMS = median(NCC_AMS1)) %>% 
  t()

wilcox.test(x = subset(full.SO, subset = type == "Strictly sorted")$NCC_AMS1,
            y = subset(full.SO, subset = type == "Leniently sorted")$NCC_AMS1)
wilcox.test( x = subset(full.SO, subset = type == "Strictly sorted")$NCC_AMS1,
             y = subset(full.SO, subset = type == "Unsorted")$NCC_AMS1)
wilcox.test( x = subset(full.SO, subset = type == "Leniently sorted")$NCC_AMS1,
             y = subset(full.SO, subset = type== "Unsorted")$NCC_AMS1)
```

# NCC markers
## Def NCCs in unsorted samp
```{r}
unsort.SO <- full.SO %>% 
  subset(cells = Cells(subset(full.SO, subset = type == "Unsorted")))

# Extract SOX10 expression (RNA assay, default data slot)
sox10_expr <- GetAssayData(unsort.SO, assay = "SCT", layer = "data")["SOX10", ]

# Create NCC classification
unsort.SO$NCC <- ifelse(sox10_expr > 0, "NCC", "Other")
rm(sox10_expr); gc()

# Optionally make it a factor with defined order
unsort.SO$NCC <- factor(unsort.SO$NCC, levels = c("Other", "NCC"))

unsort.SO$NCC %>% table()
```

## Plot unsort samp in umap
```{r}
Idents(unsort.SO) <- unsort.SO$type
P03 <- DimPlot(unsort.SO, 
               reduction = "umap_harmony", group.by = "NCC",
               order = c("NCC", "Other"),
               alpha = 0.7, stroke.size = 0, pt.size = 1) + 
  scale_color_manual(values = c("Other" = "#BDBDBD",
                                "NCC"   = "#ffc64fff")) + 
  labs(y = "umap 2", x = "umap 1") +
  theme(plot.margin = margin(t=1.7, r=1, b=1.7, l=0.4, unit = "mm"),
        axis.text.y = element_text(size = 8, angle = 45),
        axis.title.y = element_text(size = 11),
          
        axis.text.x = element_text(size = 8, angle = 45),
        axis.title.x = element_text(size = 11),

        legend.text = element_text(size = 11),
        legend.key.size = unit(2.5, "mm"),
        legend.title = element_text(size = 11),
        legend.position = "inside",
        
        plot.title = element_blank())
P03

ggsave(paste0(path, "03_Results/UMAP_Unsort.svg"),
       P03, height = 89, width = 89, units = "mm")
```

```{r}
P03_inset <- FetchData(unsort.SO, 
          vars = c("SOX10", "nCount_RNA"),
          assay = "RNA,", 
          layer = "count") %>%
  mutate(Sox10_PM = SOX10 / (nCount_RNA / 1e6),
         logSox10_PM = log1p(Sox10_PM)) %>%
  ggplot(aes(x = logSox10_PM, fill = SOX10 == 0)) +
  geom_histogram(binwidth = 1) +
  
  scale_x_continuous(breaks = log1p(c(0, 200, 800)),
                     labels = c("0", "200", "800")) +
  
  scale_y_break(breaks = c(150, 10450)) +
  scale_y_continuous(breaks = c(0, 100, 10500, 10600),
                     labels = c("0", "100", "10500", "10600"),
                     limits = c(0, 10610)) +
  
  scale_fill_manual(values = c(`TRUE`  = "#BDBDBD",   # bar at 0
                               `FALSE` = "#ffc64fff"),  # non-zero bars
                    guide = "none") +
  #labs(x = "Sox10 per 1M",  y = "Freq.") +
  theme_classic() +
  theme(plot.margin = margin(t=0, r=0, b=0, l=0),

        axis.text.y.right  = element_blank(),
        axis.ticks.y.right = element_blank(),
        axis.line.y.right = element_blank(),
        
        axis.title = element_blank(),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 45, vjust = 0.5))

ggsave(paste0(path, "03_Results/UMAP_Unsort_hist.svg"),
       P03_inset, height = 35, width = 40, units = "mm")
P03_inset
```

## Calc diff. exp.
This does differential expression testing between NCCs (Sox10 positive cells) 
and all other cells in the unsorted samples.
```{r}
Idents(unsort.SO) <- unsort.SO$NCC
DEG.df <- FindMarkers(unsort.SO, 
                          ident.1 = "NCC",
                          only.pos = FALSE,
                          logfc.threshold = 0,
                          min.pct = 0)
DEG.df$gene <- rownames(DEG.df)

#Add average normalized expression in NCCs to the dataframe
avg_NCC_exp <- unsort.SO %>%
  subset(subset = NCC == "NCC") %>%
  GetAssayData(assay = "SCT", layer = "counts") %>%
  rowMeans()
DEG.df <- DEG.df %>% mutate(avg_NCC_exp = avg_NCC_exp[rownames(DEG.df)])
rm(avg_NCC_exp); gc()

# Add negative log10 p column
DEG.df$neg_log10p <- log10(DEG.df$p_val)*-1

DEG.df %>% head()
DEG.df %>% str()
```

## Def NCC markers
```{r}
#Subset DE table
NCCmrkrs.df <- DEG.df %>%       #Use only:
  subset(avg_log2FC > 0.1) %>%  # -Genes with relevant effect size
  subset(p_val_adj <= 0.05) %>% # -Significantly different genes
  subset(avg_NCC_exp > 0.05) #%>%    # -Genes with informative mean exp in NCCs
  #subset(pct.1 >= 0.01) 

#Add column to label the NCC markers in DEG.df
DEG.df$NCCmrkrs <- ifelse(DEG.df$gene %in% NCCmrkrs.df$gene, "Yes", "No")

##Save table for supplementary as .csv
#DEG.df %>% 
#  #Select the rellevant columns
#  select(gene, avg_log2FC, avg_NCC_exp, p_val_adj, NCCmrkrs) %>%
#  #Reorder the rows to make it easier to find the markers at a glance
#  arrange(factor(NCCmrkrs, levels = c("Yes", "No")),
#          p_val_adj,
#          desc(avg_log2FC)
#          ) %>% 
#  # Round off all values 
#  mutate(avg_log2FC = round(avg_log2FC, 3),
#         avg_NCC_exp = round(avg_NCC_exp, 3),
#         p_val_adj = case_when(
#           p_val_adj < 0.001 ~ "<0.001",
#           p_val_adj > 0.001  ~ as.character(round(p_val_adj, 3)))) %>% 
#  # rename the columns for readability
#  rename("Gene"                       = gene,
#         "Average log2-fold change"   = avg_log2FC,
#         "Average SCT counts in NCCs" = avg_NCC_exp,
#         "Adjusted p-value"           = p_val_adj,
#         "NCC markers"                = NCCmrkrs) %>%
#  write.csv(paste0(path, "03_Results/",
#                   "DiffExpAll_", Sys.Date(), ".csv"), 
#            row.names = FALSE)

```

## Plotting
Here i plot differential expression in a few different ways for one of the 
figures in the paper.

## MA plot
```{r}
#Plot
P04 <- DEG.df %>%
  mutate(avg_log2FC = ifelse(avg_log2FC > 5, 5, avg_log2FC)) %>%
  mutate(avg_log2FC = ifelse(avg_log2FC < -5, -5, avg_log2FC)) %>%
  mutate(avg_NCC_exp = ifelse(avg_NCC_exp > 10, 10, avg_NCC_exp)) %>%
  arrange(NCCmrkrs == "Yes") %>% 
  
  ggplot(aes(y = avg_log2FC, x = avg_NCC_exp, color = NCCmrkrs)) +
  geom_point(shape = 19, size = .7, alpha = .7) +
  scale_x_continuous(trans = scales::log1p_trans()) +
  scale_color_manual(values = c("Yes" = "#ffc64fff", "No" = "#BDBDBD")) +
  ylab(label = TeX('$log_{2}$(fold change)')) +
  xlab(label = "Mean expressin in putative NCCs, log(x+1)")+
  labs(title = "Differential expression", 
       color = "NCC markers") +
  theme(plot.title = element_blank(),   
        axis.title.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key.size = unit(2.5, "mm"),
        legend.position = "none",
        #Background
        plot.background = element_rect(fill = "white", color = "white"),
        panel.background = element_rect(fill = "white", color = "white"),
        panel.border = element_rect(colour = "black", fill = NA)) +
  geom_vline(xintercept = .05, linetype = "dashed") + 
  geom_hline(yintercept = .2, linetype = "dashed") 
P04

ggsave(filename = paste0(path, "03_Results/MAPlot.svg"),
      plot = P04, height = 89, width = 89, units = "mm")
```

## Volcano plot
### Lebels
```{r}
# Manually set some selected genes to label
Volcano_labels <- c("SOX10", "EDNRB", "TFAP2A", "TFAP2B", "ZEB2", 
                    "PAX7", "ETS1", "SOX8", "SOX5", 
                    
                    "GYPC", "ACBD7", "RAB19")

Volcano_labels <- NCCmrkrs.df[Volcano_labels, ]

Volcano_labels <- Volcano_labels %>% 
  mutate(avg_log2FC = ifelse(avg_log2FC > 7, 7, avg_log2FC)) %>%
  mutate(avg_log2FC = ifelse(avg_log2FC < -7, -7, avg_log2FC)) %>%
  mutate(neg_log10p = ifelse(neg_log10p > 100, 100, neg_log10p))

Volcano_labels
```

### Plot
```{r}
# Set line for p value
sig_line <- DEG.df %>% 
  arrange(NCCmrkrs == "Yes") %>% {log10(0.05/nrow(.))*-1}

#Construct the plot
P05 <- DEG.df %>% 
  mutate(avg_log2FC = ifelse(avg_log2FC > 7, 7, avg_log2FC)) %>%
  mutate(avg_log2FC = ifelse(avg_log2FC < -7, -7, avg_log2FC)) %>%
  mutate(neg_log10p = ifelse(neg_log10p > 100, 100, neg_log10p)) %>%
  arrange(NCCmrkrs == "Yes") %>% 
  
  ggplot(aes(x = avg_log2FC, y = neg_log10p, color = NCCmrkrs)) + 
  geom_point(shape = 19, size = .7, alpha = .7) +
  geom_vline(xintercept = 0.2, linetype = "dashed") +
  geom_hline(yintercept = sig_line, linetype = "dashed") +
  scale_color_manual(values = c("Yes" = "#ffc64fff", "No" = "#BDBDBD")) +
  ylab(label = TeX('$-log_{10}$(adj. p-value)')) +
  xlab(label = TeX('$log_{2}$(fold change)')) +
  labs(title = "NCC markers", 
       color = "NCC markers") +
  geom_label_repel(data = Volcano_labels, 
                  #Italic lowercase gene names 
                  aes(label = paste0("italic(", str_to_sentence(gene), ")")),
                  parse = TRUE,
                  color = "black", 
                  max.overlaps = 20,
                  nudge_x = -4,
                  size = 3,
                  label.padding = 0.1,
                  label.r = .1) +
  theme(plot.title = element_blank(),   
        axis.title.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        #Background
        plot.background = element_rect(fill = "white", color = "white"),
        panel.background = element_rect(fill = "white", color = "white"),
        panel.border = element_rect(colour = "black", fill = NA),
        legend.position = "none")
P05
ggsave(filename = paste0(path, "03_Results/VolcanoPlot.svg"),
      plot = P05, height = 89, width = 89, units = "mm")
```


```{r}
P05 + geom_point(
    data = DEG.df %>% filter(gene == "GYPC"), # Change gene to label the point 
    aes(x = avg_log2FC, y = neg_log10p),       # of your choice
    color = "red", size = 2)
```

# Module score new NCC markers
```{r}
full.SO <- AddModuleScore(full.SO, 
                          list(NCCmrkrs.df$gene),
                          #Or, excluding Sox10:
                          #list(NCCmrkrs.df$gene[2:length(NCCmrkrs.df$gene)]), 
                          name = "NewNCCmrkrsAMD")
```

```{r}
P06 <- full.SO@meta.data %>%
  ggplot(aes(x = type, fill = type, y = NewNCCmrkrsAMD1)) +
  geom_violin() +
  scale_fill_manual(values = c("Unsorted"          = "#BDBDBD",
                               "Leniently sorted"  = "#53AC90",
                               "Strictly sorted"   = "#64EC09")) +
  labs(y = "NCC marker module score") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 11),
        axis.text = element_text(size = 8),
        legend.position = "none")
    
ggsave(filename = paste0(path, "03_Results/NCCmrkrVln.svg"),
      plot = P06, height = 89, width = 89, units = "mm")
P06
```

# Save data
Save the filtered data and meta data in an RDS file to be read by the next script in the pipeline (04_IntAnalysis10XAndSmrtS.Rmd)
```{r}
full.SO
saveRDS(full.SO, paste0(path, 
                       "02_Analysis/04_IntAnalysis_output_", 
                       Sys.Date(),
                       ".Rds"))
```

Calculated runtime:
```{r}
end_time <- Sys.time()
paste0("Elapsed time: ", end_time - start_time)
```
